---
title: "Condition of Lake Erie Largemouth Bass Sampled in the ODOW Nearshore Community Survey 2013-2017"
author: "Alex Benecke"
date: "January 4, 2018"
output: pdf_document
header-includes:
    - \setlength\parindent{24pt}
    - \usepackage{indentfirst}
editor_options: 
  chunk_output_type: console
---

### Data Prep

```{r setup and load packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FSA)
library(car) ## befor dplyr reduce conflicts with MASS
library(dplyr)
library(plotrix)
library(multcomp)
library(tidyverse)
library(magrittr)
library(Matching)
```

```{r Load Data, results='hide'}
Stock <- read.csv("Data/Clean-Data/largemouth-bass_Wr_Stock.csv") %>%
  filter(Year < 2017) %>%
  filterD(!is.na(Wr)) %>%
  arrange(Year,gcat)


Stock$fyr <- factor(Stock$fyr)
Stock$Year <- factor(Stock$Year)
```
```{r View Data, tidy=TRUE}
headtail(Stock)
str(Stock)
unique(Stock$fyr)
```

  **Note**
  
  *I removed the years 2012 and 2017. 2012 because only large fish have weight length data and more and differenct sites were samples. I removed 2017 due to differences in the survey.*
  
  **Note**
  
  *I am removing a fish from site 15 year 2013 because it appears to be a outlier (Wr > 200). Probably due to data entry error.* **I went back and did this in 'Create-Wr-Gabelhouse-Data.Rmd' where I make the data file I use for this analysis.**
  
  
### Summarize Relative Weight by Year

```{r Summarizing Wr Data for 2012,2014,2015,2016, tidy=TRUE}
(Wr.Stock <- Summarize(Wr~Year,data = Stock) %>%
  arrange(Year))
```


**Note**

  *The relative weight data contains only stock length individuals. This is so that I can easily compare the relative weight of fish with PSD. This is done despite the min TL being 150 mm. I may want to summarize relative weight for 150mm and greater length individuals in the future to see if young/small fish drive down or increase Wr.*


Lets start exploring the relative weight data. I have two questions I would like to know the answer to.

1) does Wr differ among years?

2) does Wr differ among gabelhouse length categories?


**First** Lets see if Wr is different between years. 

```{r Anova Wr Year, tidy=TRUE}
aov1 <- lm(Wr~Year,data=Stock)
#save(aov1,file = "model-output/aov1.rda")
Anova(aov1)
summary(aov1)

boxplot(Wr~Year,data=Stock)
```
\newpage
```{r Tukey,tidy=TRUE}
mc1 <- glht(aov1,mcp(Year="Tukey"))
summary(mc1)
```


  Looks like Wr *is significantly different* between years (One-Way ANOVA, $F_{3, 406}$ = 4.00, p = 0.007933). There is *no significant difference* in relative weight between
  2013 and 2014 (Tukey HSD, t = -1.94, p = 0.2101),
  2013 and 2015 (Tukey HSD, t = -1.52, p = 0.4241),
  2013 and 2016 (Tukey HSD, t = 1.02, p = 0.7346), 
  2015 and 2014 (Tukey HSD, t = 0.10, p = 0.9996), and 
  2015 and 2016 (Tukey HSD, t = 2.47, p = 0.0656). However, relative weight *is significantly different* between 
  2014 and 2016 (Tukey HSD, t = 3.11, p = 0.0107).


### constructing a plot of Wr and Year

```{r Create Results to be plotted, tidy=TRUE}
grps.1 <- c("2013", "2014", "2015", "2016")
nd.1 <- data.frame(Year=factor(grps.1,levels = grps.1))
(pred.1 <- predict(aov1,nd.1,interval = "confidence"))
```

```{r Plot Wr by Year, tidy=TRUE}
plotCI(as.numeric(nd.1$Year), pred.1[,"fit"],
       li = pred.1[,"lwr"], ui = pred.1[,"upr"],
       pch=19, xaxt = "n", 
       xlim = c(0.8, 4.2), ylim = c(79,121),
       xlab = "Year", ylab = "Mean Wr")

lines(nd.1$Year, pred.1[,"fit"],col="gray50")

axis(1,at=nd.1$Year, labels = nd.1$Year)

cld(mc1)

text(x=nd.1$Year, y=pred.1[,"upr"], labels = c("ab", "a", "ab", "b"), pos=3)

abline(h=100,lty=4,col="green")
```


```{r Testing Assumptions, tidy=TRUE}
residPlot(aov1)
```
\newpage
```{r, tidy=TRUE}
leveneTest(aov1) # Pitential outlier line 43
# 2013 site 15 Weight = 1714 Length = 368 Wr > 200
```

  Variance are equal and the homoscedasticity assumption is likley met (Levene's Test, $F_{3, 406}$ = 1.5423, p = 0.203). 

\vspace{0.5in}

**Without 2017** 
  *Variance are equal and the homoscedasticity assumption is likley met (Levene's Test, $F_{2, 311}$ = 1.75, p = 0.18).*


```{r real quick make Wr.csv from pred, tidy=TRUE}
Year <- c("2013", "2014", "2015", "2016")
pred.1 <- data.frame(Year,pred.1)
names(pred.1) <- c("Year", "Wr", "LCI", "UCI")
str(pred.1)
head(pred.1)
#2-9-2018#write.csv(pred.1,file = "Data/Clean-Data/summary-data/relative-weight_largemouth-bass_STOCK.csv",row.names = FALSE)
```


# 2) Wr and Gcat

  I will look into the difference in Wr between gcat at a later date. I don't think this matters so much as of now.


```{r Summarize Wr by Year and Gcat, echo=FALSE, tidy=TRUE}
Wr.gcat <- Summarize(Wr~Year + gcat,data = Stock) %>%
  arrange(gcat)

display.Wr <- data.frame(Wr.gcat[,c(1:3)],
                round(Wr.gcat[,c(4:10)],2))

print(display.Wr,row.names=FALSE)
```

**IDK this doesn't look quite right below**

```{r Anova Wr gcat Year, tidy=TRUE}
gcat.aov <- aov(Wr~  Year*gcat -1, data = Stock)
#3-7-2018#save(gcat.aov,file = "model-output/gcat.aov.rda")

anova(gcat.aov)
#summary(lm(Wr~  Year*gcat -1, data = Stock))
```

```{r gcat.aov Tukey, echo=FALSE, tidy=TRUE}
Tukey.Wr <- TukeyHSD(gcat.aov)

Tukey.Wr.gcat <- round(TukeyHSD(gcat.aov)$'Year:gcat',3)

(btwn.yr.Tky.Wr.gcat <- Tukey.Wr.gcat[c(1:3,12:13,22,
                                    39:41,46:47,52,
                                    61:66),])
```

  There is *no significant difference* in mean Wr within gabelhouse length categories among years years ($F_{6, 398}$ = 1.53, p = 0.17).

```{r boxplot Wr gcat year}

par(mar=c(6.75,4,0.25,0.25))

boxplot(Wr ~ gcat + Year, data=Stock,
        at = c(1,2,3,5,6,7,9,10,11,13,14,15),
        xaxt="n", yaxt="n")

abline(h=100,lty=2,col="green")

mtext(c("Preferred", "Quality", "Stock","Preferred", "Quality", "Stock","Preferred", "Quality", "Stock","Preferred", "Quality", "Stock"),
     side = 1,
     line = 0.25,
     las=2,
     at = c(1,2,3,5,6,7,9,10,11,13,14,15),
     cex = 1.25)

mtext(c("2013", "2014", "2015", "2016"),
     side = 1,
     line = 5,
     at = c(2,6,10,14),
     cex = 1.25)

axis(2,
     at = seq(0,150,20),
     las = 2)
```

### Wr between year within gcat

```{r}
Wr.P <- filterD(Stock,gcat=="preferred")

Wr.Q <- filterD(Stock,gcat=="quality")

Wr.S <- filterD(Stock,gcat=="stock")
```


**Preferred**
```{r}
WrP.aov <- lm(Wr ~ Year, data = Wr.P)
anova(WrP.aov)
```
```{r}
WrP.mc <- glht(WrP.aov,mcp(Year="Tukey"))
summary(WrP.mc)
```

**Quality**
```{r}
WrQ.aov <- lm(Wr ~ Year, data = Wr.Q)
anova(WrQ.aov)
```
```{r}
WrQ.mc <- glht(WrQ.aov,mcp(Year="Tukey"))
summary(WrQ.mc)
```

**Stock**
```{r}
WrS.aov <- lm(Wr ~ Year, data = Wr.S)
anova(WrS.aov)
```
```{r}
WrS.mc <- glht(WrS.aov,mcp(Year="Tukey"))
summary(WrS.mc)
```




### Wr between gcat within year
  I want to see if Wr is different between gcat within each year. Above I show where the relationships of mean Wr for particular gcats between years. 

  
```{r Seperationg out years}
Wr.13 <- filterD(Stock,Year==2013)
Wr.14 <- filterD(Stock,Year==2014)
Wr.15 <- filterD(Stock,Year==2015)
Wr.16 <- filterD(Stock,Year==2016)

str(Wr.13)
```
  
  
**2013**  
```{r}
Wr13.aov <- lm(Wr ~gcat, data = Wr.13)
anova(Wr13.aov)
```
```{r}
Wr13.mc <- glht(Wr13.aov,mcp(gcat="Tukey"))
summary(Wr13.mc)
```
  
  There is a *Significant* difference in mean Wr between gabelhouse length categories during **2013** ($F_{2,94}$ = 15.14, p < 0.001). Mean Wr of **Stock** length largemouth bass is *significantly* different from **quality** (Tukey HSD, t = 4.93, p < 0.001) and **preferred** (Tukey HSD, t = 4.11, p < 0.001) length largemouth bass.
  
  
**2014**  
```{r}
Wr14.aov <- lm(Wr ~gcat, data = Wr.14)
anova(Wr14.aov)
```
```{r}
Wr14.mc <- glht(Wr14.aov,mcp(gcat="Tukey"))
summary(Wr14.mc)
```

  There is a *Significant* difference in mean Wr between gabelhouse length categories during **2014** ($F_{2,137}$ = 26.32, p < 0.001). Mean Wr of **Stock** length largemouth bass is *significantly* different from **quality** (Tukey HSD, t = 6.04, p < 0.001) and **preferred** (Tukey HSD, t = 5.74, p < 0.001) length largemouth bass.
  
  
**2015**  
```{r}
Wr15.aov <- lm(Wr ~gcat, data = Wr.15)
anova(Wr15.aov)
```
```{r}
Wr15.mc <- glht(Wr15.aov,mcp(gcat="Tukey"))
summary(Wr15.mc)
```

  There is a *Significant* difference in mean Wr between gabelhouse length categories during **2015** ($F_{2,64}$ = 17.32, p < 0.001). Mean Wr of **Stock** length largemouth bass is *significantly* different from **quality** (Tukey HSD, t = 3.91, p < 0.001) and **preferred** (Tukey HSD, t = 4.74, p < 0.001) length largemouth bass as well as **quality** and **preferred** (Tukey HSD, t = , p = 0.007) length largemouth bass.
  
  
**2016**  
```{r}
Wr16.aov <- lm(Wr ~gcat, data = Wr.16)
anova(Wr16.aov)
```
```{r}
Wr16.mc <- glht(Wr16.aov,mcp(gcat="Tukey"))
summary(Wr16.mc)
```
  
  There is a *Significant* difference in mean Wr between gabelhouse length categories during **2016** ($F_{2,103}$ = 8.20, p < 0.001). Mean Wr of **Stock** length largemouth bass is *significantly* different from **quality** (Tukey HSD, t = 3.53, p = 0.002) and **preferred** (Tukey HSD, t = 2.87, p = 0.013) length largemouth bass.
  
  
  
  
