---
title: "Age-Length Key"
author: "Alex J. Benecke"
date: "January 31, 2017"
output: pdf_document
header-includes:  
    - \setlength\parindent{24pt}
    - \usepackage{indentfirst}
editor_options: 
  chunk_output_type: console
---

 Age-length keys will be produced from the measured TL and the age-at-capture of each Largemouth Bass. All samples will be divided into 25 mm bins?
size bins (denoted $L_{i}$) and grouped by age ($A_{j}$). The probability that a fish is a particular age given its size ($p_{j|i}$) will be calculated by dividing the number of fish ($n_{ji}$) in the $i^th$ length interval of the $j^th$ age by the total number of fish in that size interval. The calculation for $p_{j|i}$ will be performed in R using the FSA, magrittr, and dplyr packages according to the methods described by Derek Ogle (2016a) (Bache and Wickham 2016, Ogle 2016b, Wickham and Francois 2016).

```{r Loading Packages,tidy=TRUE}
library(FSA)
library(magrittr)
library(dplyr)
library(nnet)
```
```{r Load Data, tidy=TRUE}
LMB <- read.csv("Data/Clean-Data/2016_largemouth-bass_clean.csv") %>%
  select(FID,Site,AgeCap,LenCap,WTg,SEXCON,Sex) 

LMB$FID <- factor(LMB$FID)
LMB$Site <- factor(LMB$Site)
LMB$SEXCON <- factor(LMB$SEXCON)
LMB$Sex <- factor(LMB$Sex)

length(LMB$FID)
LMB[LMB$FID==55,]
#LMB <- LMB[-55,] ## Remove FID 55
length(LMB$FID)
LMB <- filterD(LMB,!is.na(FID))

str(LMB)
headtail(LMB)
```

```{r Create Length intervals, tidy=TRUE}
LMB %<>% mutate(lencat25=lencat(LenCap,w=25))
headtail(LMB)
```
```{r Seperate Aged and Unaged (NA) Fish,tidy=TRUE}
is.na(LMB$AgeCap)

LMB.unaged <- filter(LMB, is.na(AgeCap))
headtail(LMB.unaged)
all(is.na(LMB.unaged$AgeCap)) # Better be True

LMB.aged <- filter(LMB,!is.na(AgeCap))
headtail(LMB.aged)
any(is.na(LMB.aged$AgeCap)) # Better be False
```

```{r Observed Age-Length Key,tidy=TRUE}
(alk.freq <- xtabs(~lencat25+AgeCap,data = LMB.aged))

rowSums(alk.freq)

alk <- prop.table(alk.freq,margin = 1)
round(alk,3)

### Some weirdness here I have a 450 mm 2 year old (FID 55 removed) and 425 mm 6 yr old while my 8 year old is 400 mm ? Check ages on some of these!!!
```

```{r Smoothed or Modeled Age-Length Key, tidy=TRUE}
LMB.mlr <- multinom(AgeCap~lencat25,data=LMB.aged,maxit=500)
```

```{r Predicted Age at Length, tidy=TRUE}
lens <- seq(100,650,25)
alk.sm <- predict(LMB.mlr,data.frame(lencat25=lens),type = "probs")
row.names(alk.sm) <- lens # for clarity
round(alk.sm,3) #for display purposes

### Much better but still double check the ages from before!!!
```

### Smoothed Age-Length Key Model
```{r Displaying the Smoothed Age Length Key Graphically, tidy=TRUE}
### Smoothed ALK Model
alkPlot(alk.sm,type = "area",showLegend = TRUE,leg.cex = 0.7,xlab = "Total Langth (mm)")

alkPlot(alk.sm,type = "bubble", xlab = "Total Length (mm)")
```

### Age-Length Key
```{r Displaying the Age Length Key Graphically , tidy=TRUE}
alkPlot(alk,type = "area",showLegend = TRUE,leg.cex = 0.7,xlab = "Total Langth (mm)")

alkPlot(alk,type = "bubble", xlab = "Total Length (mm)")
### Doesn't Look Good!!!
```

