---
title: "Length Frequency"
author: "Alex J. Benecke"
date: "January 26, 2018"
output: pdf_document
header-includes:
    - \setlength\parindent{24pt}
    - \usepackage{indentfirst}
editor_options: 
  chunk_output_type: console
---

## Data Preparation for Length-Frequency Analysis  2013 - 2017 

  I will now compare the length frequency distribution for largemouth bass obtained in the nearshore electrofishing survey during 2013 - 2017.

  
```{r load required packages, include=FALSE}
library(FSA)
library(magrittr)
library(dplyr)
library(plotrix)
library(Matching)
```


```{r Load Data 2013 - 2017, tidy=TRUE}
lmb <- read.csv("Data/Clean-Data/2012-2017_nearshore-survey-largemouth-bass_CLEAN.csv") %>%
  arrange(Year,FID,Length) %>%
  filterD(!is.na(Length))
lmb$fyr <- as.factor(lmb$fyr)
```
```{r View Data 2013 - 2017, include=FALSE}
str(lmb)
headtail(lmb)
```
```{r Check Years 2013 - 2017}
unique(lmb$Year) ### See that there is no 2013
```

  Lets create a new variable for 20 mm length bins.

```{r lcat20 2013 - 2017, tidy=TRUE}
lmb %<>% mutate(lcat20=lencat(Length,w=20))
```
```{r View Length Bins 2013 - 2017, include=FALSE}
headtail(lmb)
```

  Now I want to separate out the years. I will throw out the year 2012 because samples from this years were not collected using the same procedures as in subsequent years. Only large LMB from 2012 had length weigh data.


```{r Separate Data by Year 2013 - 2017, tidy=TRUE}
lmb.12 <- filter(lmb,Year==2012)
#1-8-2018#write.csv(lmb.12, file = "Data/Clean-Data/minor-data/lmb.12.csv")
lmb.13 <- filter(lmb,Year==2013)
#1-8-2018#write.csv(lmb.13, file = "Data/Clean-Data/minor-data/lmb.13.csv")
lmb.14 <- filter(lmb,Year==2014)
#1-8-2018#write.csv(lmb.14, file = "Data/Clean-Data/minor-data/lmb.14.csv")
lmb.15 <- filter(lmb,Year==2015)
#1-8-2018#write.csv(lmb.15, file = "Data/Clean-Data/minor-data/lmb.15.csv")
lmb.16 <- filter(lmb,Year==2016)
#1-8-2018#write.csv(lmb.16, file = "Data/Clean-Data/minor-data/lmb.16.csv")
lmb.17 <- filter(lmb,Year==2017)
#1-8-2018#write.csv(lmb.17, file = "Data/Clean-Data/minor-data/lmb.17.csv")
```
\newpage

### Length Frequency Distribution ###


```{r Length Frequency Tables 2013 - 2017, tidy=TRUE, include=FALSE}
### 2013
(Freq13.20 <- xtabs(~lcat20,data=lmb.13))
prop.table(Freq13.20)*100

### 2014
(Freq14.20 <- xtabs(~lcat20,data=lmb.14))
prop.table(Freq14.20)*100

### 2015
(Freq15.20 <- xtabs(~lcat20,data=lmb.15))
prop.table(Freq15.20)*100

### 2016
(Freq16.20 <- xtabs(~lcat20,data=lmb.16))
prop.table(Freq16.20)*100

### 2017
(Freq17.20 <- xtabs(~lcat20,data=lmb.17))
prop.table(Freq17.20)*100
```

  Lets view a quick histogram of the frequency of fish in each length bin.


```{r Hist of Length Freq 2013 - 2016, cache=TRUE, tidy=TRUE, echo=FALSE}
par(mfrow=c(5,1)) ### arrange graphs 1 column 4 rows
par(mfrow=c(2,2))
par(mfrow=c(1,1)) ### reset to default

### 2013
hist(~Length,data=lmb.13,
     breaks = seq(0,500,20),
     ylim=c(0,30),
     main="2013",
     xlab="Total Length (mm)")

### 2014
hist(~Length,data=lmb.14,
     breaks = seq(0,500,20),
     ylim=c(0,30),
     main="2014",
     xlab="Total Length (mm)")

### 2015
hist(~Length,data=lmb.15,
     breaks = seq(0,500,20),
     ylim=c(0,30),
     main="2015",
     xlab="Total Length (mm)")

### 2016
hist(~Length,data=lmb.16,
     breaks = seq(0,500,20),
     ylim=c(0,30),
     main="2016",
     xlab="Total Length (mm)")

### 2017
hist(~Length,data=lmb.17,
     breaks = seq(0,500,20),
     ylim=c(0,30),
     main="2017",
     xlab="Total Length (mm)")
```

  I can tell that small fish (<100-250mm) are definitely not being sampled consistently. **2014** and **2016** look more or less as I would expect with a stable largemouth bass population (2014 more so than 2016) with a peak and more or less gradual decline in the number of fish with increasing length. The peak for **2015** appears to be shifted right (320mm) more so than for 2014 and 2016 (260mm). The year **2013** Appears to be more or less flat but it's highest peak is at (360mm) before declining. As for **2017** I think this represents inadequate sampling I see a peak with much smaller fish (100mm) than other years and some length bins with only 1 fish. the peak for 2017 is around 340mm.
  
  I would have to say from these graphs that the length frequency distribution of largemouth bass is unstable. However, I am tempted to think that there may be some aspect of sampling bias playing a role in this. **2017** was not sampled fully or by the same gear or crew (University of Toledo and old shock boat) due to staff shortages so I think I may throw this year out (missing 4 annual sites but has 2 rotating sites). **2013** is also missing several sites (6, 12, 19) although one site is NA so this could be one of those three. However, **2013** does seem to have a good sample size (n = 114) also site 19 was only sampled in 2017 and site 6 never seemed to provide alot of largemouth bass. So I will keep the year 2013 in at least for now. **2015** was also incompletely sampled missing two annual sotes (10 and 19) but with one rotating site for a total of nine sites. **2015** has a lower sample size than other years (n=80) *except for 2017*. However, since I dont't have any years in which all sites were sampled  and the same number of sites were sampled in 2015 as in 2014 and 2016 I cannot throw out this year. Although, just looking at the data from 2015 it seems for whatever reason we have a significantly smaller proportion of smaller (stock and substock) fish.

\newpage

## Length Frequency 2013 - 2016 


### Cumulative Frequencies 2013 - 2016 ###

  Lets look at the empirical cumulative distribution function (ECDF). This is the proportion of fish less than each observed length. This should help me compare the length frequency distributions between years.


```{r ECDF 2013 - 2016, cache=TRUE, tidy=TRUE, echo=FALSE}
par(mfrow=c(1,1)) ### reset to default

### 2013
plot(ecdf(lmb.13$Length),
     xlab="Total Length (mm)",
     do.points=FALSE,
     verticals=TRUE, 
     main="",
     col.01line = NULL, 
     col="orange")
### 2014
plot(ecdf(lmb.14$Length),
     xlab="Total Length (mm)", 
     do.points=FALSE,
     verticals=TRUE, 
     main="", 
     col.01line = NULL,
     col="red",
     add=T)
### 2015
plot(ecdf(lmb.15$Length),
     xlab="Total Length (mm)", 
     do.points=FALSE,
     verticals=TRUE, 
     main="", 
     col.01line = NULL,
     col="blue", add=T)
### 2016
plot(ecdf(lmb.16$Length),
     xlab="Total Length (mm)",
     do.points=FALSE,
     verticals=TRUE, main="",
     col.01line = NULL, 
     col="green",
     add=T)

legend("bottomright", 
       c("2013","2014","2015","2016"),
       col=c("orange","red","blue","green"),
       lty=1, bty="n", cex=0.75)
```

### Compare Length Frequency Between Years 2013 - 2016 ###
#### Kolmogorov-Smirnov Test ####

  I will test wether the cumulative frequencies are different between years 2013 - 2016 using a Kolmogorov-Smirnov Two-tailed test. An example of this test that I used to help me teach myself how to do this can be found ( https://ned.ipac.caltech.edu/level5/Wall2/Wal4_3.html ). The formula I used to calculate the D value that would be significant at a given level of significance can be found ( https://ned.ipac.caltech.edu/level5/Wall2/TableA_9.html ).

```{r Calculating D for given level of Significance, echo=FALSE,results="hide"}
(yrs <- c("13-14","13-15","13-16","14-15","14-16","15-16"))
sqrt.mn <- function(a,b){
  m = length(a)
  n = length(b)
  result = sqrt((m + n)/(m*n))
  result
}

(tmp.x <- c(sqrt.mn(lmb.13$Length,lmb.14$Length),
           sqrt.mn(lmb.13$Length,lmb.15$Length),
           sqrt.mn(lmb.13$Length,lmb.16$Length),
           sqrt.mn(lmb.14$Length,lmb.15$Length),
           sqrt.mn(lmb.14$Length,lmb.16$Length),
           sqrt.mn(lmb.15$Length,lmb.16$Length)))
(df.tmp <- data.frame(yrs,tmp.x))

df.tmp

(p0.05 <- 1.36)
(p0.01 <- 1.63)
(p0.005 <- 1.73)
(p0.001 <- 1.95)

D.0.05 <- numeric(length(df.tmp$tmp.x))

D.0.01 <- numeric(length(df.tmp$tmp.x))

D.0.005 <- numeric(length(df.tmp$tmp.x))
  
D.0.001 <- numeric(length(df.tmp$tmp.x))

for(i in 1:length(df.tmp$tmp.x)){
  D.0.05[i] = p0.05 * (df.tmp$tmp.x[i])
  D.0.05
  
  D.0.01[i] = p0.01 * (df.tmp$tmp.x[i])
  D.0.01
  
  D.0.005[i] <- p0.005*df.tmp$tmp.x[i]
  D.0.005
  
  D.0.001[i] <- p0.001*df.tmp$tmp.x[i]
  D.0.001
  
  result2 <- data.frame(D.0.05,D.0.01,D.0.005,D.0.001)
  result2
}

(df.tmp2 <- data.frame(df.tmp,result2))


```

|     Year    |    m   |   n  |     0.05     |      0.01     |     0.005      |     0.001      |
|:-----------:|:------:|:----:|:------------:|:-------------:|:--------------:|:--------------:|
| 2013 - 2014 |   114  | 143  |  0.1707596   |   0.2046604   |   0.2172162    |   0.2448391    |
| 2013 - 2015 |   114  |  80  |  0.1983546   |   0.2377339   |   0.2523188    |   0.2844055    |
| 2013 - 2016 |   114  | 131  |  0.1741943   |   0.2087770   |   0.2215854    |   0.2497638    |
| 2014 - 2015 |   143  |  80  |  0.1898796   |   0.2275763   |   0.2415380    |   0.2722539    |
| 2014 - 2016 |   143  | 131  |  0.1644790   |   0.1971329   |   0.2092270    |   0.2358339    |
| 2015 - 2016 |    80  | 131  |  0.1929743   |   0.2312854   |   0.2454747    |   0.2766911    |


```{r Kolmogorov-Smirnov Test 2013 - 2016, cache=TRUE, tidy=TRUE, include=FALSE}
### 2013 and 2014
(ks.13.14 <- ks.test(lmb.13$Length,lmb.14$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties
ksb.13.14 <- ks.boot(lmb.13$Length,lmb.14$Length,nboots = 5000)
summary(ksb.13.14)

### 2013 and 2015
(ks.13.15 <- ks.test(lmb.13$Length,lmb.15$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties
ksb.13.15 <- ks.boot(lmb.13$Length,lmb.15$Length,nboots = 5000)
summary(ksb.13.15)

### 2013 and 2016
(ks.13.16 <- ks.test(lmb.13$Length,lmb.16$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties
ksb.13.16 <- ks.boot(lmb.13$Length,lmb.16$Length,nboots = 5000)
summary(ksb.13.16)

### 2014 and 2015
(ks.14.15 <- ks.test(lmb.14$Length,lmb.15$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties

ksb.14.15 <- ks.boot(lmb.14$Length,lmb.15$Length,nboots = 5000)
summary(ksb.14.15)

### 2014 and 2016
(ks.14.16 <- ks.test(lmb.14$Length,lmb.16$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties

ksb.14.16 <- ks.boot(lmb.14$Length,lmb.16$Length,nboots = 5000)
summary(ksb.14.16)

### 2015 and 2016
(ks.15.16 <- ks.test(lmb.15$Length,lmb.16$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties

ksb.15.16 <- ks.boot(lmb.15$Length,lmb.16$Length,nboots = 5000)
summary(ksb.15.16)


### Adjusted p-values

(ks.ps <- c(ks.test(lmb.13$Length,lmb.14$Length)$p.value,
            ks.test(lmb.13$Length,lmb.15$Length)$p.value,
            ks.test(lmb.13$Length,lmb.16$Length)$p.value,
            ks.test(lmb.14$Length,lmb.15$Length)$p.value,
            ks.test(lmb.14$Length,lmb.16$Length)$p.value,
            ks.test(lmb.15$Length,lmb.16$Length)$p.value))
(p.val <- p.adjust(ks.ps))
```
```{r Make Adjuster p-values a little easier to read 2013 - 2016, results="hide", cache=TRUE, tidy=TRUE, echo = FALSE}
(D <- c(ks.13.14$statistic[[1]], 
        ks.13.15$statistic[[1]],
        ks.13.16$statistic[[1]],
        ks.14.15$statistic[[1]], 
        ks.14.16$statistic[[1]],
        ks.15.16$statistic[[1]]))
(yrs <- c("13-14","13-15","13-16","14-15","14-16","15-16"))
```
```{r Lets look at results from the K-S test 2013 - 2016, echo=FALSE, results="hide"}
(p.yr <- data.frame(yrs,D,p.val))


(tmp.3 <- data.frame(Years = yrs,
                     p.value = round(p.yr$p.val,5),
                     D = round(D,5),
                     D.at.0.05 = round(df.tmp2$D.0.05,5),
                     D.at.0.01 = round(df.tmp2$D.0.01,5),
                     D.at.0.005 = round(df.tmp2$D.0.005,5),
                     D.at.0.001 = round(df.tmp2$D.0.001,5)))

```

\newpage

### Summary of Results 2013 - 2016 ###

  The results of the Kolmogorov-Smirnov test above seem to suggest the largemouth bass population is not stable (*Or is it? I think there are just a few weird years probably sampling related*). The length frequency distribution **is significant different** between the years 2013 and 2014 (D = 0.24, p < 0.05), 2014 and 2016 (D = 0.25, p < 0.05), and 2015 and 2016 (D = 0.24, p < 0.05). There is **no significant difference** between the length frequency distributions for 2013 and 2015 (D = 0.20, p = 0.15), 2013 and 2016 (D = 0.10, p = 0.58), and 2014 and 2015 (D = 0.19, p = 0.15).
  
  
|     Year    |      D     |    D at p of 0.05     |   p-value   | Relationship |
|:-----------:|:----------:|:---------------------:|:-----------:|:------------:|
| 2013 - 2014 |   0.23721  |       0.17076         |    0.008    |       S      |
| 2013 - 2015 |   0.19518  |       0.19835         |    0.148    |      NS      |
| 2013 - 2016 |   0.09997  |       0.17419         |    0.576    |      NS      |
| 2014 - 2015 |   0.19003  |       0.18988         |    0.148    |      NS      |
| 2014 - 2016 |   0.24748  |       0.16448         |    0.003    |       S      |
| 2015 - 2016 |   0.24265  |       0.19297         |    0.023    |       S      |

http://www.introspective-mode.org/kolmogorov-smirnov-goodness-of-fit/



## Length Frequency 2013 - 2016 Stock Length (>= 200mm)

  *What is I remove small fish (<200mm) which I am not sure are fully sampled by my gear. Will 2014 still be different from 2013 and 2016? will 2015?*

  
| Year | n Sub-Stock |  n Stock  |  n Quality  |  n Preferred  |
|:----:|:-----------:|:---------:|:-----------:|:-------------:|
| 2013 |     16      |    41     |    41       |      16       |
| 2014 |      3      |    65     |    57       |      18       |
| 2015 |     13      |    14     |    38       |      15       |
| 2016 |     24      |    52     |    44       |      11       |

  
  **2014** has a lot of *stock* length individuals ($n_{stock} = 65$) compared to **2013** ($n_{stock} = 41$), **2015** ($n_{stock} = 14$), and **2016** ($n_{stock} = 52$) but far fewer *substock* individuals ($n_{sun-stock} = 3$), **2013** ($n_{sun-stock} = 16$), **2015** ($n_{sun-stock} = 13$), **2016** ($n_{sun-stock} = 24$). **2014** also seems to have a bit more but comperable amount of *quality* and *preferred* length individuals (57,18 respectivley) than **2013** (41,16), **2015** (38,15), and **2016** (44,11). So, lets explore wether the differences are in the smaller sub stock length not fully sampled by our gear.
  
  
```{r Load Data Stock,, echo=FALSE, results="hide", tidy=TRUE}
lmb.s <- read.csv("Data/Clean-Data/2012-2016_nearshore-survey-largemouth-bass_Stock_CLEAN.csv") %>%
  arrange(Year,FID,Length)
lmb.s$fyr <- as.factor(lmb.s$fyr)
```
```{r view Stock data, include=FALSE}
str(lmb.s)
headtail(lmb.s)
```
```{r check years Stock, echo=FALSE, results="hide"}
unique(lmb.s$Year) ### See that there is no 2013
table(lmb.s$gcat) ### Check that there are no sub-stock fish
```
```{r lcat20 Stock, echo=FALSE, tidy=TRUE}
lmb.s %<>% mutate(lcat20=lencat(Length,w=20))
```
```{r Stock, include=FALSE}
headtail(lmb.s)
```
```{r new df by year Stock, echo=FALSE, tidy=TRUE}
lmbs.12 <- filter(lmb.s,Year==2012)
lmbs.13 <- filter(lmb.s,Year==2013)
lmbs.14 <- filter(lmb.s,Year==2014)
lmbs.15 <- filter(lmb.s,Year==2015)
lmbs.16 <- filter(lmb.s,Year==2016)
```


```{r Calc m and n for Stock Length Fish, echo=FALSE, results="hide"}
stck.mn <- function(a,b){
  m = a
  n = b
  result = sqrt((m + n)/(m*n))
  result
}

(tmp.s <- c(stck.mn(98,140),
           stck.mn(98,67),
           stck.mn(98,107),
           stck.mn(140,67),
           stck.mn(140,107),
           stck.mn(67,107)))
(df.stck <- data.frame(yrs,tmp.s))

df.stck


Ds.0.05 <- numeric(length(df.stck$tmp.s))


for(i in 1:length(df.stck$tmp.s)){
  Ds.0.05[i] = p0.05 * (df.stck$tmp.s[i])
  Ds.0.05
}

(df.stck2 <- data.frame(df.stck,Ds.0.05))
```


|     Year    |    m   |   n  |     0.05     |  
|:-----------:|:------:|:----:|:------------:|
| 2013 - 2014 |    98  | 140  |  0.1317078   |
| 2013 - 2015 |    98  |  67  |  0.1585227   |
| 2013 - 2016 |    98  | 107  |  0.1398209   |
| 2014 - 2015 |   140  |  67  |  0.1485538   |
| 2014 - 2016 |   140  | 107  |  0.1284081   |
| 2015 - 2016 |    67  | 107  |  0.1557921   |

\newpage

### Length Frequency Distribution of Stock Length Fish During 2013 - 2016 ###

```{r length frequency tables Stock, tidy=TRUE, include=FALSE}
### 2013
(Freq13s.20 <- xtabs(~lcat20,data=lmbs.13))
prop.table(Freq13s.20)*100 ## Peak at 360 mm

### 2014
(Freq14s.20 <- xtabs(~lcat20,data=lmbs.14))
prop.table(Freq14s.20)*100 ## Peak at 260mm

### 2015
(Freq15s.20 <- xtabs(~lcat20,data=lmbs.15))
prop.table(Freq15s.20)*100 ## Peak at 320mm

### 2016
(Freq16s.20 <- xtabs(~lcat20,data=lmbs.16))
prop.table(Freq16s.20)*100 ## Peak at 260mm 
```

  Lets view a quick histogram of the frequency of fish in each length bin.

```{r Hist of length freq Stock, cache=TRUE, tidy=TRUE, echo=FALSE}
par(mfrow=c(5,1)) ### arrange graphs 1 column 4 rows
par(mfrow=c(2,2))
par(mfrow=c(1,1)) ### reset to default

### 2013
hist(~Length,data=lmbs.13,
     breaks = seq(180,500,20),
     ylim=c(0,30),
     main="2013",
     xlab="Total Length (mm)")

### 2014
hist(~Length,data=lmbs.14,
     breaks = seq(180,500,20),
     ylim=c(0,30),
     main="2014",
     xlab="Total Length (mm)")

### 2015
hist(~Length,data=lmbs.15,
     breaks = seq(180,500,20),
     ylim=c(0,30),
     main="2015",
     xlab="Total Length (mm)")

### 2016
hist(~Length,data=lmbs.16,
     breaks = seq(180,500,20),
     ylim=c(0,30),
     main="2016",
     xlab="Total Length (mm)")
```

\newpage

#### Cumulative Frequencies of Stock Length Fish Only During 2013 - 2016 ####

  Lets look at the empirical cumulative distribution function (ECDF). This is the proportion of fish less than each observed length. This should help me compare the length frequency distributions between years.

```{r ECDF Stock, cache=TRUE, tidy=TRUE, echo=FALSE}
par(mfrow=c(1,1)) ### reset to default

### 2013
plot(ecdf(lmbs.13$Length),
     xlab="Total Length (mm)",
     do.points=FALSE,
     verticals=TRUE, 
     main="",
     col.01line = NULL, 
     col="orange")
### 2014
plot(ecdf(lmbs.14$Length),
     xlab="Total Length (mm)", 
     do.points=FALSE,
     verticals=TRUE, 
     main="", 
     col.01line = NULL,
     col="red",
     add=T)
### 2015
plot(ecdf(lmbs.15$Length),
     xlab="Total Length (mm)", 
     do.points=FALSE,
     verticals=TRUE, 
     main="", 
     col.01line = NULL,
     col="blue", add=T)
### 2016
plot(ecdf(lmbs.16$Length),
     xlab="Total Length (mm)",
     do.points=FALSE,
     verticals=TRUE, main="",
     col.01line = NULL, 
     col="green",
     add=T)

legend("bottomright", 
       c("2013","2014","2015","2016"),
       col=c("orange","red","blue","green"),
       lty=1, bty="n", cex=0.75)
```

### Compare Length Frequency Between Years 2013 - 2016 Stock Length Fish Only ###
#### Kolmogorov-Smirnov Test Of Only Stock Length Fish ####


```{r Kolmogorov-Smirnov Test Stock, cache=TRUE, tidy=TRUE, include=FALSE}
### 2013 and 2014
(ks.13.14s <- ks.test(lmbs.13$Length,lmbs.14$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties
ksb.13.14s <- ks.boot(lmbs.13$Length,lmbs.14$Length,nboots = 5000)
summary(ksb.13.14s)

### 2013 and 2015
(ks.13.15s <- ks.test(lmbs.13$Length,lmbs.15$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties
ksb.13.15s <- ks.boot(lmbs.13$Length,lmbs.15$Length,nboots = 5000)
summary(ksb.13.15s)

### 2013 and 2016
(ks.13.16s <- ks.test(lmbs.13$Length,lmbs.16$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties
ksb.13.16s <- ks.boot(lmbs.13$Length,lmbs.16$Length,nboots = 5000)
summary(ksb.13.16s)

### 2014 and 2015
(ks.14.15s <- ks.test(lmbs.14$Length,lmbs.15$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties

ksb.14.15s <- ks.boot(lmbs.14$Length,lmbs.15$Length,nboots = 5000)
summary(ksb.14.15s)

### 2014 and 2016
(ks.14.16s <- ks.test(lmbs.14$Length,lmbs.16$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties

ksb.14.16s <- ks.boot(lmbs.14$Length,lmbs.16$Length,nboots = 5000)
summary(ksb.14.16s)

### 2015 and 2016
(ks.15.16s <- ks.test(lmbs.15$Length,lmbs.16$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties

ksb.15.16s <- ks.boot(lmbs.15$Length,lmbs.16$Length,nboots = 5000)
summary(ksb.15.16s)


### Adjusted p-values

(ks.ps.s <- c(ks.test(lmbs.13$Length,lmbs.14$Length)$p.value,
            ks.test(lmbs.13$Length,lmbs.15$Length)$p.value,
            ks.test(lmbs.13$Length,lmbs.16$Length)$p.value,
            ks.test(lmbs.14$Length,lmbs.15$Length)$p.value,
            ks.test(lmbs.14$Length,lmbs.16$Length)$p.value,
            ks.test(lmbs.15$Length,lmbs.16$Length)$p.value))
(p.val.s <- p.adjust(ks.ps.s))
```
```{r Make Adjuster p-values a little easier to read Stock, cache=TRUE, tidy=TRUE, echo = FALSE}
(D.s <- c(ks.13.14s$statistic[[1]], 
        ks.13.15s$statistic[[1]],
        ks.13.16s$statistic[[1]],
        ks.14.15s$statistic[[1]], 
        ks.14.16s$statistic[[1]],
        ks.15.16s$statistic[[1]]))
(yrs <- c("13-14","13-15","13-16","14-15","14-16","15-16"))
```
```{r Lets look at results from the K-S test Stock, echo=FALSE, results="hide"}
(p.yr.s <- data.frame(Years = yrs,
                      D = round(D.s,5),
                      D.at.0.05 = round(df.stck2[,3],5),
                      p.value = p.val.s))
```

### Summary of Results 2013 - 2016 Stock Length Fish Only ###

   The length frequency distribution **is significant different** between the years 
   2014 and 2015 (D = 0.27, p < 0.05), and 
   2015 and 2016 (D = 0.28, p < 0.05). There is **no significant difference** between the length frequency distributions for 
   2013 and 2014 (D = 0.15, p = 0.46), 
   2013 and 2015 (D = 0.25, P = 0.06), 
   2013 and 2016 (D = 0.09, P = 0.98), and 
   2014 and 2016 (D = 0.11, P = 0.98).
  
  
|     Year    |      D     | D at alpha of 0.05 |  p-value   |  Relationship |
|:-----------:|:----------:|:------------------:|:----------:|:-------------:|
| 2013 - 2014 |   0.14898  |      0.17912       |   0.46     |      NS       |
| 2013 - 2015 |   0.24627  |      0.21559       |   0.06     |      ~S       |
| 2013 - 2016 |   0.08974  |      0.19016       |   0.98     |      NS       |
| 2014 - 2015 |   0.26962  |      0.20203       |   0.02     |       S       |
| 2014 - 2016 |   0.10721  |      0.17464       |   0.98     |      NS       |
| 2015 - 2016 |   0.27703  |      0.21188       |   0.02     |       S       |


  
## 2013 - 2017 Length Frequency

#### Cumulative Frequencies 2013 - 2017 ####

  Lets look at the empirical cumulative distribution function (ECDF). This is the proportion of fish less than each observed length. This should help me compare the length frequency distributions between years.


```{r ECDF 2013 - 2017, cache=TRUE, tidy=TRUE, echo=FALSE}
par(mfrow=c(1,1)) ### reset to default

### 2013
plot(ecdf(lmb.13$Length),
     xlab="Total Length (mm)",
     do.points=FALSE,
     verticals=TRUE, 
     main="",
     col.01line = NULL, 
     col="orange")
### 2014
plot(ecdf(lmb.14$Length),
     xlab="Total Length (mm)", 
     do.points=FALSE,
     verticals=TRUE, 
     main="", 
     col.01line = NULL,
     col="red",
     add=T)
### 2015
plot(ecdf(lmb.15$Length),
     xlab="Total Length (mm)", 
     do.points=FALSE,
     verticals=TRUE, 
     main="", 
     col.01line = NULL,
     col="blue", add=T)
### 2016
plot(ecdf(lmb.16$Length),
     xlab="Total Length (mm)",
     do.points=FALSE,
     verticals=TRUE, main="",
     col.01line = NULL, 
     col="green",
     add=T)
### 2017
plot(ecdf(lmb.17$Length),
     xlab="Total Length (mm)",
     do.points=FALSE,
     verticals=TRUE, main="",
     col.01line = NULL, 
     col="purple",
     add=T)

legend("bottomright", 
       c("2013","2014","2015","2016","2017"),
       col=c("orange","red","blue","green","purple"),
       lty=1, bty="n", cex=0.75)
```

```{r Kolmogorov-Smirnov Test 2013 - 2017, cache=TRUE, tidy=TRUE, include=FALSE}
### 2013 and 2014
(ks.13.14 <- ks.test(lmb.13$Length,lmb.14$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties
ksb.13.14 <- ks.boot(lmb.13$Length,lmb.14$Length,nboots = 5000)
summary(ksb.13.14)

### 2013 and 2015
(ks.13.15 <- ks.test(lmb.13$Length,lmb.15$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties
ksb.13.15 <- ks.boot(lmb.13$Length,lmb.15$Length,nboots = 5000)
summary(ksb.13.15)

### 2013 and 2016
(ks.13.16 <- ks.test(lmb.13$Length,lmb.16$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties
ksb.13.16 <- ks.boot(lmb.13$Length,lmb.16$Length,nboots = 5000)
summary(ksb.13.16)

### 2013 and 2017
(ks.13.17 <- ks.test(lmb.13$Length,lmb.17$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties
ksb.13.17 <- ks.boot(lmb.13$Length,lmb.17$Length,nboots = 5000)
summary(ksb.13.17)

### 2014 and 2015
(ks.14.15 <- ks.test(lmb.14$Length,lmb.15$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties

ksb.14.15 <- ks.boot(lmb.14$Length,lmb.15$Length,nboots = 5000)
summary(ksb.14.15)

### 2014 and 2016
(ks.14.16 <- ks.test(lmb.14$Length,lmb.16$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties

ksb.14.16 <- ks.boot(lmb.14$Length,lmb.16$Length,nboots = 5000)
summary(ksb.14.16)

### 2014 and 2017
(ks.14.17 <- ks.test(lmb.14$Length,lmb.17$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties

ksb.14.17 <- ks.boot(lmb.14$Length,lmb.17$Length,nboots = 5000)
summary(ksb.14.17)

### 2015 and 2016
(ks.15.16 <- ks.test(lmb.15$Length,lmb.16$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.15$Length) :
#   p-value will be approximate in the presence of ties

ksb.15.16 <- ks.boot(lmb.15$Length,lmb.16$Length,nboots = 5000)
summary(ksb.15.16)

### 2015 and 2017
(ks.15.17 <- ks.test(lmb.15$Length,lmb.17$Length))
# Warning message:
# In ks.test(lmb.14$Length, lmb.17$Length) :
#   p-value will be approximate in the presence of ties

ksb.15.17 <- ks.boot(lmb.15$Length,lmb.17$Length,nboots = 5000)
summary(ksb.15.17)

### 2016 and 2017
(ks.16.17 <- ks.test(lmb.16$Length,lmb.17$Length))
# Warning message:
# In ks.test(lmb.16$Length, lmb.17$Length) :
#   p-value will be approximate in the presence of ties

ksb.16.17 <- ks.boot(lmb.16$Length,lmb.17$Length,nboots = 5000)
summary(ksb.16.17)

### Adjusted p-values

(ks.ps <- c(ks.test(lmb.13$Length,lmb.14$Length)$p.value,
            ks.test(lmb.13$Length,lmb.15$Length)$p.value,
            ks.test(lmb.13$Length,lmb.16$Length)$p.value,
            ks.test(lmb.13$Length,lmb.17$Length)$p.value,
            ks.test(lmb.14$Length,lmb.15$Length)$p.value,
            ks.test(lmb.14$Length,lmb.16$Length)$p.value,
            ks.test(lmb.14$Length,lmb.17$Length)$p.value,
            ks.test(lmb.15$Length,lmb.16$Length)$p.value,
            ks.test(lmb.15$Length,lmb.17$Length)$p.value,
            ks.test(lmb.16$Length,lmb.17$Length)$p.value))
(p.val <- p.adjust(ks.ps))
```
```{r Make Adjuster p-values a little easier to read 2013 - 2017, cache=TRUE, tidy=TRUE, echo = FALSE, results="hide"}
(D <- c(ks.13.14$statistic[[1]], 
        ks.13.15$statistic[[1]],
        ks.13.16$statistic[[1]],
        ks.13.17$statistic[[1]],
        ks.14.15$statistic[[1]], 
        ks.14.16$statistic[[1]],
        ks.14.17$statistic[[1]],
        ks.15.16$statistic[[1]],
        ks.15.17$statistic[[1]],
        ks.16.17$statistic[[1]]))
(yrs <- c("13-14","13-15","13-16","13-17","14-15","14-16","14-17","15-16","15-17","16-17"))
```
```{r Lets look at results from the K-S test 2013 - 2017, echo=FALSE, results="hide"}
(p.yr <- data.frame(yrs,D,p.val))
```

### Summary of Results 2013 - 2017 ###

  The results of the Kolmogorov-Smirnov test above seem to suggest the largemouth bass population is not stable (*Or is it? I think there are just a few weird years probably sampling related*). The length frequency distribution **is significant different** between the years 2013 and 2014 (D = 0.24, P = 0.014), 2014 and 2016 (D = 0.25, P < 0.005), and 2015 and 2016 (D = 0.24, P = 0.046). There is **no significant differnce** between the length frequency distributions for 2013 and 2015 (D = 0.20, P = 0.344), 2013 and 2016 (D = 0.10, P = 1), and 2014 and 2015 (D = 0.19, P = 0.344). The length frequency distribution for the year 2017 was *not significantly different* between any years. 

**Note:**
*Adding in the Year 2017 Significantly Altered the Adjusted P-Values*

  
|     Year    |    D    |    p-value   | Relationship |
|:-----------:|:-------:|:------------:|:------------:|
| 2013 - 2014 |   0.24  |     0.014    |      D       |
| 2013 - 2015 |   0.20  |     0.344    |      S       |
| 2013 - 2016 |   0.10  |     1.000    |      S       |
| 2013 - 2017 |   0.14  |     1.000    |      S       |
| 2014 - 2015 |   0.19  |     0.344    |      S       |
| 2014 - 2016 |   0.25  |     0.005    |      D       |
| 2014 - 2017 |   0.20  |     0.686    |      S       |
| 2015 - 2016 |   0.24  |     0.046    |      D       |
| 2015 - 2017 |   0.14  |     1.000    |      S       |
| 2016 - 2017 |   0.17  |     1.000    |      S       |

  
