---
title: "Proportional Size Distribution (PSD)"
author: "Alex J. Benecke"
date: "December 19, 2017"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

  I will use a Chi-Square test to see if there is a difference between PSD between years.

```{r load required packages}
library(FSA)
library(magrittr)
library(dplyr)
library(plotrix)
library(Matching)
```


```{r Load Data}
lmbs <- read.csv("Data/Clean-Data/2012-2016_nearshore-survey-largemouth-bass_Stock_CLEAN.csv") %>%
  filter(Year>=2013) %>%
  arrange(Year,FID,Length)
lmbs$fyr <- as.factor(lmbs$fyr)

str(lmbs)
headtail(lmbs)
unique(lmbs$Year) ### See that there is no 2013
```

```{r Create a Table of Fish in each gcat by Year}
(lmbs.LF <- xtabs(~Year+gcat,data=lmbs))
```


```{r Chi-Square test}
chisq.test(lmbs.LF)
```

  This seems to suggest that the distribution of fish into the length intervals is different between years ($X^{2}$ = 16.694, df = 6, P = 0.01048).
  
  Lets try and find where the differences may be.
  
```{r percent of fish in each gcat by year}
round(prop.table(lmbs.LF,margin=1)*100,0)
```

  Remarkably the percent of quality fish is the same for 2014 and 2016 and a bit higher for 2015. the percentage of fish in each gcat is similar between years 2013, 2014, and 2016, However the year 2015 appears to have a higher percentage of large fish and far fewer small fish. 
  
  Could this be some sort of sampling bias?
  
  Could this be a result of sampling different sites? where the 2015 sites more suitable for LMB?
  
  Are the years really different or do I just have too few to say for sure?

















