---
title: "Growth Figures"
author: "Alex J. Benecke"
date: "January 22, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FSA)
library(tidyverse)
library(nlme)
```
```{r Load and Prep Data, tidy=TRUE}
LMBL <- read.csv("Data/Clean-Data/2016_largemouth-bass_long-format.csv") %>%
  arrange(FID,Agei)

### Making factors factors
LMBL$FID <- factor(LMBL$FID)
LMBL$Site <- factor(LMBL$Site)
LMBL$SEXCON <- factor(LMBL$SEXCON)
LMBL$Sex <- factor(LMBL$Sex)
```
```{r Display Data, eval=FALSE}
str(LMBL)
headtail(LMBL)
```
Creating the von Bertalanffy function.
```{r Build von Bertalanffy Growth Model, tidy=TRUE, eval=FALSE}
LVB <- function(x, Linf, K, t0){
  y = Linf * (1 - exp(-K * (x - t0)))
  y
}
LVB <- vbFuns()

LVB(5, 422.8, 0.39, -0.40)
LVB(5, Linf = c(422.8, 0.39, -0.40)) ### Should be the same output
```
#### Removing the individual with no sex (FID=89)
```{r remove no sex individual, tidy=TRUE, eval=FALSE}
### Just looking at data
head(LMBL)
### Finding fish with unknown sex
(unknown.sex <- filterD(LMBL,Sex==0))
### Getting row number for fish with the unknown sex
(FID89 <- as.numeric(row.names(LMBL[LMBL$Sex==0,])))
### removing the fish with unknown sex from the data set
length(LMBL$FID) ### just seeing the number of rows in the data set
length(unique(LMBL$FID)) ### just seeing the number of fish
LMBL <- LMBL[-c(FID89),]
length(LMBL$FID)
length(unique(LMBL$FID)) ### Good! looks like only FID 89 was removed
### Lets make sure there is no empty row in my data 
LMBL <- filterD(LMBL,!is.na(FID))
### and lets just take a quick look at the data
str(LMBL)
headtail(LMBL)
```
#### Remake Grouped Data Object
```{r Group Data Without Unknown Sex Fish}

datgr = groupedData(BI.len ~ Agei|FID, data = LMBL,
                  labels = list(x = "Age", y = "Size"),
                  units = list(x = "(Years)", y = "(mm)"))
```








```{r Load Model Output, tidy=TRUE}
### The outputs commented out failed to converge
load("model-output/nlme.mod2.rda")

#load("model-output/sexmod.lkt.rda")
#load("model-output/sexmod.lk.rda")
load("model-output/sexmod.lt.rda")
load("model-output/sexmod.kt.rda")
load("model-output/sexmod.l.rda")
load("model-output/sexmod.k.rda")
load("model-output/sexmod.t.rda")

#load("model-output/sexmod.lkt.int.rda")
#load("model-output/sexmod.lk.int.rda")
#load("model-output/sexmod.lt.int.rda")
#load("model-output/sexmod.kt.int.rda")
load("model-output/sexmod.l.int.rda")
#load("model-output/sexmod.k.int.rda")
#load("model-output/sexmod.t.int.rda")
```



### Results of {$L_{\infty}$}, Sex Model, {sexmod.l}

```{r Results of sexmod.l, echo=FALSE, tidy=TRUE}
summary(sexmod.l)
intervals(sexmod.l)
```

```{r sexmod.l Plot and effects, include=FALSE}
fixef(sexmod.l)
ranef(sexmod.l)
coef(sexmod.l)[1,3]
```

par(cex = 2,)
las=2 always perpendicular to axis
lwd =
mai =c(b,l,t,r) margin in inches
mar
A numerical vector of the form c(bottom, left, top, right) which gives the number of lines of margin to be specified on the four sides of the plot. The default is c(5, 4, 4, 2) + 0.1.

tck
The length of tick marks as a fraction of the smaller of the width or height of the plotting region. If tck >= 0.5 it is interpreted as a fraction of the relevant side, so if tck = 1 grid lines are drawn. The default setting (tck = NA) is to use tcl = -0.5.
```{r Prepare to Plot sexmod.l, echo=FALSE, tidy=TRUE}
Axes <- seq(100,500,by=50)
Years <- seq(0,10,by=1)

## plot individual fish data
## plot the fixed parameter model
## plot individual fish models

### rgb(0,1,0,0.25, maxColorValue=1)

par(mar=c(3.25,3.25,0,0),mgp=c(2.1,0.4,0),tcl=-0.2,
    lwd=2,
    cex=1.25,
    cex.lab = 1.25,
    cex.axis = 1)
```


```{r plot sexmod.l}
plot(jitter(LMBL$Agei),LMBL$BI.len,
     col=ifelse(LMBL$Sex==1,rgb(0,0,1,0.4, maxColorValue=1),rgb(1,0,0,0.4, maxColorValue=1)),
     pch=19,
     ylim=c(100,500),
     xlim=c(0,10),
     xlab = "Age (Years)",
     ylab = "Back-Calculated Length (mm)",
     bty="n",
     yaxt="n",
     xaxt="n")

axis(2,at = Axes, las=2, lwd=2)
axis(1,at = Years, las=1,lwd=2)

abline(h=425.4478748,lty=2,col="blue") ### Males
abline(h=439.0623179,lty=2,col="red") ### females

abline(h=457,lty=4,col="green")

x <- seq(1,11,by=1)
lines(x, fixef(sexmod.l)[1] * (1 - exp(-fixef(sexmod.l)[3] * (x - (fixef(sexmod.l)[4])))),
      lwd=3, 
      col="blue") ### Males

lines(x, fixef(sexmod.l)[2] * (1 - exp(-fixef(sexmod.l)[3] * (x - (fixef(sexmod.l)[4])))),
      lwd=3, 
      col="red") ### Females
```
I think I need to do a for loop with an if else statement for male and female.
lets see if this works. it failed. I need to figure out which object I can assign male and females to for the if statemens (i.e., sexmod.l$??? == 1).

```{r load Male and Female model output}
tmp.male <- read.csv("Data/Clean-Data/tmp.male.csv")
headtail(tmp.male)

tmp.female <- read.csv("Data/Clean-Data/tmp.female.csv")
headtail(tmp.female)
```

```{r Males}
head(tmp.male);nrow(tmp.male)

for(i in 1:56){
  lines(x, (tmp.male[i,2] + tmp.male[i,6]) * (1 - exp(- tmp.male[i,4] * (x - (tmp.male[i,5])))),
        lwd=3,
        col=rgb(0,0,1,0.15),
        lty=3) } ### Males

```


```{r Females}
head(tmp.female);nrow(tmp.female)

for(i in 1:69){
  lines(x, (tmp.female[i,3] + tmp.female[i,6]) * (1 - exp(- tmp.female[i,4] * (x - (tmp.female[i,5])))),
        lwd=3,
        col=rgb(0.85,0,0,0.15),
        lty=3) } ### Females
```

```{r Legend}
legend("topleft",
       legend = print(expression(L[i]==425 %*% (1 - e **{-0.39 %*% (t[i] + 0.39)}))),
       bty="n",
       cex=1.5,
       text.col = "blue") ### VB Equation Males

legend("topright",
       legend = print(expression(L[i]==439 %*% (1 - e **{-0.39 %*% (t[i] + 0.39)}))),
       bty="n",
       cex=1.5,
       text.col = rgb(0.85,0,0)) ### VB Equation Females

legend("bottomright",
       legend = c("Male", "Female"),
       col=c(rgb(0,0,1), rgb(0.85,0,0)),
       lty=1, pch=19, bty="n", cex=1.5)
```







###############################


```{r Males}
#length(unique(LMBL$FID[LMBL$Sex==1])) ### n = 56 males

for(i in 1:56){
lines(x, (coef(sexmod.l)[i,1] + coef(sexmod.l)[i,5]) * (1 - exp(- coef(sexmod.l)[i,3]   
* (x - ( coef(sexmod.l)[i,4] )))),lwd=3,col=rgb(0,0,1,0.15),
lty=3) } ### Males

```
```{r Females}
#length(unique(LMBL$FID[LMBL$Sex==2])) ### n = 69 females

for(i in 1:69){
lines(x, (coef(sexmod.l)[i,2] + coef(sexmod.l)[i,5]) * (1 - exp(- coef(sexmod.l)[i,3]   
* (x - ( coef(sexmod.l)[i,4] )))),lwd=3,col=rgb(0.85,0,0,0.15),
lty=3) } ### Females
```


```{r both sexes}
for(i in 1:125){
  if(sexmod.l$plist$Linf$fixed[,1]==1){
    lines(x, (coef(sexmod.l)[i,1] + coef(sexmod.l)[i,5]) * (1 - exp(- coef(sexmod.l)[i,3]   
* (x - ( coef(sexmod.l)[i,4] )))),lwd=3,col=rgb(0,0,1,0.1),lty=3)  ### Males
  } else{
    lines(x, (coef(sexmod.l)[i,2] + coef(sexmod.l)[i,5]) * (1 - exp(- coef(sexmod.l)[i,3]   
* (x - ( coef(sexmod.l)[i,4] )))),lwd=3,col=rgb(1,0,0,0.1),lty=3)  ### Females
  }
}
```


####################


```{r Old second part of vonB graph}
for(i in 1:125){
lines(x, coef(sexmod.l)[i,1] * (1 - exp(- coef(sexmod.l)[i,3]   
* (x - ( coef(sexmod.l)[i,4] )))),lwd=3,col=rgb(0,0,1,0.1),lty=3) } ### Males

lines(x, fixef(sexmod.l)[2] * (1 - exp(-fixef(sexmod.l)[3] * (x - (fixef(sexmod.l)[4])))),
      lwd=3, 
      col="red") ### Females

for(i in 1:125){
lines(x, coef(sexmod.l)[i,2] * (1 - exp(- coef(sexmod.l)[i,3]   
* (x - ( coef(sexmod.l)[i,4] )))),lwd=3,col=rgb(1,0,0,0.1),lty=3) } ### Females

legend("topleft",
       legend = print(expression(L[i]==425 %*% (1 - e **{-0.39 %*% (t[i] + 0.39)}))),
       bty="n",
       cex=1.15,
       text.col = "blue") ### VB Equation Males

legend("topright",
       legend = print(expression(L[i]==439 %*% (1 - e **{-0.39 %*% (t[i] + 0.39)}))),
       bty="n",
       cex=1.15,
       text.col = "red") ### VB Equation Females

legend("bottomright",
       legend = c("Male", "Female"),
       col=c("blue", "red"),
       lty=1, pch=19, bty="n")
```










