---
title: "Total Mortality"
author: "Alex J. Benecke"
date: "February 8, 2017"
output: pdf_document
header-includes:
    - \setlength\parindent{24pt}
    - \usepackage{indentfirst}
editor_options: 
  chunk_output_type: console
---

see smith et al. 2012 

```{r Load Packages, include=FALSE, tidy=TRUE}
library(FSA)
library(car)
library(dplyr)
library(magrittr)
library(Rcapture)
```
```{r Load Data, include=FALSE, tidy=TRUE}
LMB <- read.csv("Data/Clean-Data/2016_largemouth-bass_clean.csv") %>%
  select(FID,Site,AgeCap,LenCap,WTg,SEXCON,Sex) %>%
  filterD(!is.na(AgeCap)) %>%
  arrange(FID)
LMB$FID <- factor(LMB$FID)
LMB$Site <- factor(LMB$Site)
LMB$SEXCON <- factor(LMB$SEXCON)
LMB$Sex <- factor(LMB$Sex)
str(LMB)
headtail(LMB)
length(LMB$FID)
LMB[LMB$FID==55,];FID55 <- as.numeric(rownames(LMB[LMB$FID==55,]))
#LMB <- LMB[-c(FID55),] %>%
#  filterD(!is.na(FID))
#length(LMB$FID)

### Create Catch df
catch <- data.frame(table(LMB$AgeCap))
names(catch) <- c(Var1 = "Age", Freq = "ct")
catch$Age <- as.numeric(catch$Age)
catch
str(catch)
```

```{r Catch per age, echo=FALSE, tidy=TRUE}
plot(log(ct)~Age,data=catch,xlab="Age (years)",ylab="log Catch",pch=19)
```


```{r Chapman-Robson Method, tidy=TRUE}
LMB.cr <- chapmanRobson(ct~Age,data=catch,ages2use=2:8)

### Show the survival and Instantaneous mortality estimates 
### from chapman-robson method
(cr.mort <- cbind(summary(LMB.cr),confint(LMB.cr)))

### Calc Annual Mortality from Z
(A.cr <- 1-exp(-coef(LMB.cr)[[2]]))
### Calc Annual Mort 95% CI
(Acr.CI <- 1-exp(-confint(LMB.cr)[2,]))
### Make nice table
(mort.A <- c(A.cr, Acr.CI)) 
names(mort.A) <- c(v1="Estimate",v2="95% LCI",v3="95% UCI")
### Show all mortality and survival estimates
cr.mort
mort.A

plot(LMB.cr,
     cex.lab=2,
     cex.est = 2,
     bty="n")

```

  Chapman-Robson $A = 0.4502858, 95\% CI 0.3209307 - 0.5550002$.

Chapman-Robson method is prefered (Dunn et al. 2002,Smith et al. 2012 {D.Ogle 2016 Book})

  Instantaneous annual mortality (Chapman-Robson Method, $Z = 0.5983568, sde = 0.1078207, LCI = 0.3870321, UCI = 0.8096815$). Annual mortality calculated from instantaneous annual mortality ($A = 0.4502858$).

  Instantaneous annual mortality ($Z$) was found to be $0.5983568$ with approximate $95$\% confidence intervals between $0.3870321$ and $0.8096815$. The estimated annual mortality rate ($A$) is $0.4502858$ with approximate $95$\% confidence intervals between $0.3209307$ and $0.5550002$. 


#### Below I explore other calculations for mortality

## Remove age 3 Yearclass 2013

```{r Chapman-Robson Method Without Age 3, tidy=TRUE}
LMB.cr_B <- chapmanRobson(ct~Age,data=catch,ages2use=c(2,4:8))
cbind(summary(LMB.cr_B),confint(LMB.cr_B))
A.crB <- 1-exp(-coef(LMB.cr_B)[[2]])
A.crB
Acr.CIB <- 1-exp(-confint(LMB.cr_B)[2,])
Acr.CIB

plot(LMB.cr_B)
```



```{r Regression Method (Single Estimate), tidy=TRUE}
tmp <- filter(catch,Age>=2) %>% mutate(lnct=log(ct))
lm1 <- lm(lnct~Age,data=tmp)
coef(lm1)
confint(lm1)
### weighted regression
tmp%<>% mutate(wts=predict(lm1))
lm2 <- lm(lnct~Age,data=tmp,weights = wts)
coef(lm2)
confint(lm2)

### Same thing but with catchCurve() from FSA
LMB.lm <- catchCurve(ct~Age,data=catch,ages2use=2:8,weighted = TRUE)
cbind(summary(LMB.lm),confint(LMB.lm))
plot(LMB.lm)
```

