---
title: "CPUE"
author: "Alex J. Benecke"
date: "February 22, 2018"
output: 
  pdf_document:
    toc: true
    toc_depth: 4
header-includes:
    - \setlength\parindent{24pt}
    - \usepackage{indentfirst}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(FSA)
library(tidyverse)
library(magrittr)
library(knitr)
```


# Hypothesis

1) $H_0$: There is no difference in cpe (catch/hour) among years 2013 - 2016.

2) $H_0$: There is no differenct in cpe for quality length (300mm) and larger largemouth bass among years 2013 - 2016.

3) $H_0$: There is no difference in cpe for largemouth bass smaller than quality length (300mm) among years 2013 - 2016.


\newpage


# 1) $H_0$: There is no difference in cpe (catch/hour) among years 2013 - 2016. 

## Load and Prepare Data ##

**Load Data**

```{r Load cpe data, tidy=TRUE}
gcat <- read.csv("Data/Clean-Data/CPUE-gcat_2013-2017.csv") %>%
  filterD(Species==317) %>%
  arrange(Site,Year)

gcat$Site <- factor(gcat$Site) 

cpeSum <- aggregate(cpe.hr ~ Year + Site, data = gcat, FUN = sum)
```
```{r, include=FALSE}
str(cpeSum)
headtail(cpeSum)
```

\centering
\begin{minipage}{0.4\textwidth}
```{r, echo=FALSE}
knitr::kable(cpeSum[c(1:21),], format = "latex")
```
\end{minipage}
\begin{minipage}{0.4\textwidth}
```{r, echo=FALSE}
knitr::kable(cpeSum[c(22:42),], format = "latex")
```
\end{minipage}
\flushleft

## Test Hypothesis 1 ##
```{r Anova mean cpe.hr and Year, tidy=TRUE}
aov1 <- aov(cpe.hr~Year,data = cpeSum)
summary(aov1)
```
```{r, echo=FALSE}
boxplot(cpe.hr~Year,data = cpeSum,
            main="Mean Catch Per Hour")
```
```{r, echo=FALSE}
(cpe.median <- aggregate(cpe.hr ~ Year, data = cpeSum, FUN = median))
```

## Results $H_0$ 1 ##

  There is *no significant difference* in CPUE between years ($F_{1,40}$ = 0.071, p = 0.792). The median CPUE was 48.36 in **2013**, 38.41 in **2014**, 22.12 in **2015**, 21.53 in **2016**. 





# 2) $H_0$: There is no differenct in cpe for quality length (300mm) and larger largemouth bass among years 2013 - 2016. 


## Load and Prepare Data ## 

**Load Data with Gcat and make Q+ and Q-**

```{r, echo=FALSE}
headtail(gcat)
```
\newpage
```{r, echo=FALSE}
xtabs(caught ~ gcat + Year, data = gcat)
```

**Make Qcat Variable and Data Frame**
```{r Make Qcat Variable, tidy=TRUE}
Qcat <- gcat %>% mutate(gcatQ=mapvalues(gcat,
                                 from=c("substock", "stock","quality","preferred", "memorable", "trophy"),
                                 to=c("quality-", "quality-", "quality+", "quality+", "quality+", "quality+"))) %>%
  dplyr::select(Year, Site, gcatQ, cpe.hr)
```

**Create Data Frame With Only Quality + Fish**

```{r Create Qpls Object, tidy=TRUE}
Qpls <- Qcat[Qcat$gcatQ=="quality+",] 
Qpls$gcatQ <- droplevels(Qpls$gcatQ)

str(Qpls)

Qpls.sum <- aggregate(cpe.hr ~ Year + Site, data = Qpls, FUN = sum) %>%
    arrange(Site,Year)
```
\centering
\begin{minipage}{0.4\textwidth}
```{r, echo=FALSE}
knitr::kable(Qpls.sum[c(1:21),], format = "latex")
```
\end{minipage}
\begin{minipage}{0.4\textwidth}
```{r, echo=FALSE}
knitr::kable(Qpls.sum[c(22:42),], format = "latex")
```
\end{minipage}
\flushleft

## Test Hypothesis 2 ##
```{r Anova of Q+ LMB, tidy=TRUE}
aov.Qpls <- aov(cpe.hr ~ Year, data = Qpls.sum)
summary(aov.Qpls)
```
```{r, echo=FALSE}
boxplot(cpe.hr ~ Year, data = Qpls.sum,
            main="Mean Catch Per Hour Quality +")
```
```{r, echo=FALSE}
(Qpls.median <- aggregate(cpe.hr ~ Year, data = Qpls.sum, FUN = median))
```


## Results $H_0$ 2 ##

  There is *no significant difference* in CPUE for fish > Quality length (300mm) among years 2013 - 2016 ($F_{1,40}$ = 0.515, p = 0.477). The median CPUE for largemouth bass Quality length and greater was 16.13 in **2013**, 22.12 in **2014**, 18.80 in **2015**, 12.64 in **2016**.






# 3) $H_0$: There is no difference in cpe for largemouth bass smaller than quality length (300mm) among years 2013 - 2016. 

## Load and Prepare Data ##

**Create Q- Data Frame**
```{r Create Q- Object, tidy=TRUE}
Qless <- Qcat[Qcat$gcatQ=="quality-",] 
Qless$gcatQ <- droplevels(Qless$gcatQ)

str(Qless)

Qless.sum <- aggregate(cpe.hr ~ Year + Site, data = Qless, FUN = sum) %>%
    arrange(Site,Year)
```
\centering
\begin{minipage}{0.4\textwidth}
```{r, echo=FALSE}
knitr::kable(Qless.sum[c(1:21),], format = "latex")
```
\end{minipage}
\begin{minipage}{0.4\textwidth}
```{r, echo=FALSE}
knitr::kable(Qless.sum[c(22:42),], format = "latex")
```
\end{minipage}
\flushleft

## Test Hypothesis 3 ##
```{r Anova Q- LMB, tidy=TRUE}
aov.Qless <- aov(cpe.hr ~ Year, data = Qless.sum)
summary(aov.Qless)
```
```{r, echo=FALSE}
boxplot(cpe.hr ~ Year, data = Qless.sum,
            main="Mean Catch Per Hour Quality-")
```
```{r, echo=FALSE}
(Qless.median <- aggregate(cpe.hr ~ Year, data = Qless.sum, FUN = median))
```

## Results $H_0$ 3 ##

  There is *no significat difference* in CPUE for fish < Quality length among years 2013 - 2016 ($F_{1,40}$ = 0.014, p = 0.906). The median CPUE for largemouth bass less than quality length was 28.85 in **2013**, 7.34 in **2014**, 4.31 in **2015**, 10.64 in **2016**.

\vspace{2em}

```{r, echo=FALSE, results='hide'}
QcatSum <- aggregate(cpe.hr ~ Year + Site + gcatQ, data = Qcat, FUN = sum) %>% 
  arrange(Site,Year,gcatQ)
QcatSum
```

```{r, echo=FALSE}
boxplot(cpe.hr ~ gcatQ*Year, data = QcatSum,
        at = c(1,2,4,5,7,8,10,11),
        col = c(rgb(0,0,1,0.75),rgb(1,0,0,0.75)),
        xaxt = "n",yaxt = "n",
        ylab = "",
        las=2,
        bty = "n")

axis(2,
     at = seq(0,140,20),
     las = 2)

mtext(c("2013", "2014", "2015", "2016"),
     side = 1,
     line = 1.5,
     at = c(1.5,4.5,7.5,10.5),
     cex = 1.25)

mtext("Year",
      side = 1,
      line = 3,
      cex = 1.5)


mtext("Mean CPUE (Hours)",
      side = 2,
      line = 2.75,
      cex = 1.5)

legend("topleft",
       legend = c("Quality +", "Quality -"),
       col = c(rgb(0,0,1,0.75),rgb(1,0,0,0.75)),
       pch = 15,
       cex = 1.5,
       bty = "n")
```

