---
title: "Condition of Lake Erie Largemouth Bass Sampled in the ODOW Nearshore Community Survey 2014-2017"
author: "Alex Benecke"
date: "January 4, 2018"
output: pdf_document
header-includes:
    - \setlength\parindent{24pt}
    - \usepackage{indentfirst}
editor_options: 
  chunk_output_type: console
---

### Data Prep

```{r setup and load packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FSA)
library(car) ## befor dplyr reduce conflicts with MASS
library(dplyr)
library(plotrix)
library(multcomp)
library(tidyverse)
library(magrittr)
library(Matching)
```

```{r Load Data, results='hide'}
Stock <- read.csv("Data/Clean-Data/2012-2017_nearshore-survey-largemouth-bass_Stock_CLEAN.csv") %>% 
  filter(fyr>=2014) %>%
  filter(!is.na(Wr)) %>%
  arrange(Year,gcat)

Stock$fyr <- as.factor(Stock$fyr)

headtail(Stock)
```
```{r View Data, tidy=TRUE}
str(Stock)
unique(Stock$fyr)
```

  **Note**
  
  *I removed the years 2012 and 2013. 2012 because only large fish have weight length data and more and differenct sites were samples. 2013 lacks weight data due to the loss of unifying ID variable and weight and lengths being seperated on different tabs.*
  
\newpage

### Summarize Relative Weight by Year

```{r Summarizing Wr Data for 2012,2014,2015,2016, tidy=TRUE}
(Wr.Stock <- Summarize(Wr~fyr,data = Stock) %>%
  arrange(fyr))
```

  I have created a file with the relative weight of each gabelhouse length category for each year. The file name is relative-weight_largemouth-bass_STOCK.csv.


**Note**

  *The relative weight data contains only stock length individuals. This is so that I can easily compare the relative weight of fish with PSD. This is done despite the min TL being 150 mm. I may want to summarize relative weight for 150mm and greater length individuals in the future to see if young/small fish drive down or increase Wr.*


Lets start exploring the relative weight data. I have two questions I would like to know the answer to.

1) does Wr differ among years?

2) does Wr differ among gabelhouse length categories?


**First** Lets see if Wr is different between years. 

```{r Anova Wr Year, tidy=TRUE}
aov1 <- lm(Wr~fyr,data=Stock)
#save(aov1,file = "model-output/aov1.rda")
Anova(aov1)

mc1 <- glht(aov1,mcp(fyr="Tukey"))
summary(mc1)
```


  Looks like Wr *is significantly different* between years (One-Way ANOVA, $F_{3, 345}$ = 7.17, p < 0.001). There is *no significant difference* in relative weight between 2015 and 2014 (Tukey HSD, t = 0.085, p = 1), 2016 and 2014 (Tukey HSD, t = 2.56, p = 0.06), 2016 and 2015 (Tukey HSD, t = 2.00, p = 0.18), and 2016 and 2017 (Tukey HSD, t = 2.37, p = 0.08). However, relative weight *is significantly different* between 2017 and 2014 (Tukey HSD, t = 4.16, p < 0.001), and 2017 and 2015 (Tukey HSD, t = 3.71, p = 0.001).


### constructing a plot of Wr and Year

```{r Create Results to be plotted, tidy=TRUE}
grps.1 <- c("2014", "2015", "2016", "2017")
nd.1 <- data.frame(fyr=factor(grps.1,levels = grps.1))
(pred.1 <- predict(aov1,nd.1,interval = "confidence"))

```

```{r Plot Wr by Year, tidy=TRUE}
plotCI(as.numeric(nd.1$fyr), pred.1[,"fit"],
       li = pred.1[,"lwr"], ui = pred.1[,"upr"],
       pch=19, xaxt = "n", 
       xlim = c(0.8, 4.2), ylim = c(65,135),
       xlab = "Year", ylab = "Mean Wr")

lines(nd.1$fyr, pred.1[,"fit"],col="gray50")

axis(1,at=nd.1$fyr, labels = nd.1$fyr)

cld(mc1)

text(x=nd.1$fyr, y=pred.1[,"upr"], labels = c("a", "a", "ab", "b"), pos=3)

abline(h=100,lty=4,col="green")
```


```{r Testing Assumptions, tidy=TRUE}
residPlot(aov1)
leveneTest(aov1)
```

  This seems to suggest that the data violates the assumption that errors are normally distributed ($F_{3,345}$ = 9.97, p < 0.001). 

  **Without 2017** *Variance are equal and the homoscedasticity assumption is likley met (Levene's Test, $F_{2, 311}$ = 1.75, p = 0.18).*


```{r real quick make Wr.csv from pred, tidy=TRUE}
Year <- c("2014", "2015", "2016", "2017")
pred.1 <- data.frame(Year,pred.1)
names(pred.1) <- c("Year", "Wr", "LCI", "UCI")
str(pred.1)
head(pred.1)
#1-10-2018#write.csv(pred.1,file = "Data/Clean-Data/relative-weight_largemouth-bass_STOCK.csv")
```

### Kruskal-Wallis Test

  A non-parametric equivalent to a one-way ANOVA. This will test if distributions are similarly shaped with equal variances among groups, that the medians are equal among all groups.
  
```{r Kruskal-Wallis Test}
kruskal.test(Wr~fyr,data = Stock)
```

  This seems to sugest that the median for at least one group differs from the medians of one or more of the years (Kruskal-Wallis rank sum test, Kruskal-wallis chi-squared = 17.5, df = 3, p < 0.001).
  
```{r Dunn Multiple Comparison}
dunnTest(Wr~fyr,data = Stock, method = "bonferroni")
dunnTest(Wr~fyr,data = Stock, method = "holm")
```

**Holm adjusted p-value**

  The Wr *is significantly different* between 2014 and 2016 (Dunn Multiple Comarison Test, Z = -3.77, p < 0.001), and 2015 and 2016 (Dunn Multiple Comarison Test, Z = -2.64, p = 0.041). However, Wr is not different between any other years.





### 2014-2016

  I want to redo the above analysis without the 2017 which I have decided not to use due to differences in data collection. I will leave the above code for now incase I or one of my coauthors wants to reference it later.


```{r Anova Wr Year 2014 - 2016, tidy=TRUE}
aov2 <- lm(Wr~fyr,data=Stock[Stock$Year<2017,])
#save(aov2,file = "model-output/aov2.rda")
Anova(aov2)

mc2 <- glht(aov2,mcp(fyr="Tukey"))
summary(mc2)
```

  Looks like relative weight *is significantly different* between years (One-Way ANOVA, $F_{2, 311}$ = 5.04, p = 0.007). There is *no significant difference* in relative weight between 2014 and 2015 (Tukey HSD, t = 0.1, p = 0.994). However, relative weight *is significantly different* between 2014 and 2016 (Tukey HSD, t = 2.98, p = 0.008), and 2015 and 2016 (Tukey HSD, t = 2.36, p = 0.049).


### constructing a plot of Wr and Year 2014 - 2016

```{r Create Results to be plotted 2014 - 2016, tidy=TRUE}
grps <- c("2014", "2015", "2016")
nd <- data.frame(fyr=factor(grps,levels = grps))
(pred <- predict(aov2,nd,interval = "confidence"))

```

```{r Plot Wr by Year 2014 - 2016, tidy=TRUE}
plotCI(as.numeric(nd$fyr), pred[,"fit"],
       li = pred[,"lwr"], ui = pred[,"upr"],
       pch=19, xaxt = "n", 
       xlim = c(0.8, 3.35), ylim = c(75,125),
       xlab = "Year", ylab = "Mean Wr")

lines(nd$fyr, pred[,"fit"],col="gray50")

axis(1,at=nd$fyr, labels = nd$fyr)

cld(mc2)

text(x=nd$fyr, y=pred[,"upr"], labels = c("a", "a", "b"), pos=3)

abline(h=100,lty=4,col="green")
```


```{r Testing Assumptions 2014 - 2016, tidy=TRUE}
residPlot(aov2)
leveneTest(aov2)
```

  Variance are equal and the homoscedasticity assumption is likley met (Levene's Test, $F_{2, 311}$ = 1.75, p = 0.18).


```{r real quick make Wr.csv from pred 2014 - 2016, tidy=TRUE}
Year <- c("2014", "2015", "2016")
pred <- data.frame(Year,pred)
names(pred) <- c("Year", "Wr", "LCI", "UCI")
str(pred)
head(pred)
#1-17-2018#write.csv(pred,file = "Data/Clean-Data/2014-16_relative-weight_largemouth-bass_STOCK.csv")
```




### To Be Continued...

  I will look into the difference in Wr between gcat at a later date. I don't think this matters so much as of now.


```{r, tidy=TRUE}
Wr.14 <- filterD(Stock,Year==2014)
Wr.15 <- filterD(Stock,Year==2015)
Wr.16 <- filterD(Stock,Year==2016)
```