---
title: "Create Data Frame with Relative Weight and Gabelhouse Length Categories"
author: "Alex J. Benecke"
date: "December 19, 2017"
output: pdf_document
header-includes:
    - \setlength\parindent{24pt}
    - \usepackage{indentfirst}
editor_options: 
  chunk_output_type: console
---

  Here I need to make two data files. Both need to cantain fish caught in the years 2012-2016 (even though I will only be using years 2014-2016). Both will contain the Relative Weight (Wr) of each fish and the gabel house length category each fish fits into. Then I will create two CLEAN data files one of wich will contain only fish larger than stock length and another with all fish of any length.
  
  
  The data file with fish of all lengths will be used to compare the length frequency distribution between years. The data frame with only gish stock length and larger will be used to compare the proportional size densities between years and with the relative weight between years.

```{r setup and load packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FSA)
library(car) ## befor dplyr reduce conflicts with MASS
library(dplyr)
library(plotrix)
library(multcomp)
library(tidyverse)
library(magrittr)
library(Matching)
```

   Now that I have some data from the 2017 nearsore survey I want to add this into the rest of the data (1-10-2018).
```{r Load 2017 Data}
LMB.17 <- read.csv("Data/Raw-Data/2017_largemouth-bass.csv") %>%
  mutate(logW=log10(Weight),logL=log10(Length))
LMB.17$fyr <- factor(LMB.17$Year)

headtail(LMB.17)
```
```{r load data}
Cond <- read.csv("./Data/Raw-Data/2012-2016_nearshore-survey_largemouth-bass.csv") %>%
  mutate(logW=log10(Weight),logL=log10(Length))
Cond$fyr <- factor(Cond$Year)

str(Cond)
headtail(Cond)
unique(Cond$Year)
```
```{r Join 2017 and the Rest of the Data}
head(LMB.17)
head(Cond) ### Just making sure I have the same variables

Cond <- rbind(Cond,LMB.17)
headtail(Cond)
unique(Cond$Year)
```



```{r Data Exploration Quick Discription, include=FALSE}
### 2013
tmp.13 <- filterD(Cond,Year==2013)
str(tmp.13)
unique(tmp.13$Year)
unique(tmp.13$Site);length(unique(tmp.13$Site)) ### Sites; Number of Sites
nrow(tmp.13) ### n Fish
mean(tmp.13$Length)

### 2014
tmp.14 <- filterD(Cond,Year==2014)
str(tmp.14)
unique(tmp.14$Year)
unique(tmp.14$Site);length(unique(tmp.14$Site)) ### Sites; Number of Sites
nrow(tmp.14) ### n Fish
mean(tmp.14$Length)


### 2015
tmp.15 <- filterD(Cond,Year==2015)
str(tmp.15)
unique(tmp.15$Year)
unique(tmp.15$Site);length(unique(tmp.15$Site)) ### Sites; Number of Sites
nrow(tmp.15) ### n Fish
mean(tmp.15$Length)


### 2016
tmp.16 <- filterD(Cond,Year==2016)
str(tmp.16)
unique(tmp.16$Year)
unique(tmp.16$Site);length(unique(tmp.16$Site)) ### Sites; Number of Sites
nrow(tmp.16) ### n Fish
mean(tmp.16$Length)

### 2017
tmp.17 <- filterD(Cond,Year==2017)
str(tmp.17)
unique(tmp.17$Year)
unique(tmp.17$Site);length(unique(tmp.17$Site)) ### Sites; Number of Sites
nrow(tmp.17) ### n Fish
mean(tmp.17$Length)
```

| Year  | Number of Fish | Number of Sites |       Unique           |
|:-----:|:--------------:|:---------------:|:----------------------:|
| 2013  |      114       |         8       | ~1 site (NA)           |
| 2014  |      143       |        11       | 2 unique (1 & 16)      |
| 2015  |       80       |         9       | 1 Unique (5)           |
| 2016  |      131       |        10       | 1 Unique (14)          |
| 2017  |       43       |         9       | 2 unique (3 & 9)       |
| Total |      511       |        18       | 6 maybe 7 Unique Sites |
  
  12 nearshore sites were sampled annually 2013 - 2017, 10 sites are sampled every year in addition to 2 sites which are sampled every five years. Length and weight data used in our analysis were obtained from 114 largemouth bass caught during 2013, 143 during 2014, 80 during 2015, and 132 during 2016. During 2013 - 2016 a total of 469 largemouth bass were obtained from 18 sites.    
  

```{r Getting Standard Weight values from literature (FSA Package)}
(wsLMB <- wsVal("Largemouth Bass", simplify = TRUE))
(wsLMB_min <- wsLMB[["min.TL"]])
(wsLMB_int <- wsLMB[["int"]])
(wsLMB_slp <- wsLMB[["slope"]])
```


```{r Calculating Standard and Relative Weight}
Cond %<>% mutate(Ws = 10^(wsLMB_int+wsLMB_slp*logL),
                 Wr=(Weight/Ws)*100)
headtail(Cond[,c(1,3,14,15)])
headtail(Cond[Cond$Year==2013,]) ### No Wr for 2013
```

### Creating data file with all size fish

```{r}
### creating size breaks for Gabelhouse Length categories for Largemouth Bass
(lmb.cuts2 <- psdVal("Largemouth Bass"))

### adding gcat variable to data frame
lmb <- Cond %>%
  mutate(gcat=lencat(Length, breaks = lmb.cuts2,
                     use.names = TRUE, drop.levels = TRUE))  ### create Gabelhouse Length Categories

  
headtail(lmb[,c(1,3,5,14:16)])
lmb[c(275:335),c(1,3,5,14:16)]
#1-10-2018#write.csv(lmb,file="Data/Clean-Data/2012-2017_nearshore-survey-largemouth-bass_CLEAN.csv") 
```


### Creating Data File with Only Stock and Larger Fish

```{r Seperating out only Stock Length individuals and removing NA}
### adding gcat variable to data frame
Stock <- Cond %>%
  filter(Length>=lmb.cuts2["stock"]) %>%
  mutate(gcat=lencat(Length, breaks = lmb.cuts2,
                     use.names = TRUE, drop.levels = TRUE))   ### create Gabelhouse Length Categories

  
headtail(Stock[,c(1,3,5,14:16)])
Stock[c(275:335),c(1,3,5,14:16)]
#1-10-2018#write.csv(Stock,file="Data/Clean-Data/2012-2017_nearshore-survey-largemouth-bass_Stock_CLEAN.csv")
```


### Creating a Data File to Summarize Relative Weight by Year for Stock Length Individuals


```{r Summarizing Wr Data for 2012,2014,2015,2016}
Stock %<>% filterD(!is.na(Wr))

Summarize(Wr~fyr, data=Stock, digits = 0) ### Wr Weight by Year

(Wr.fyr.gcat_Stock <- Summarize(Wr~fyr*gcat, data=Stock) %>%
    arrange(fyr,gcat))
str(Wr.fyr.gcat_Stock)

#1-10-2018#write.csv(Wr.fyr.gcat_Stock,file = "Data/Raw-Data/relative-weight_largemouth-bass_STOCK_RAW.csv")
```

  I have created a file with the relative weight of each gabelhouse length category for each year. The file name is relative-weight_largemouth-bass_RAW.csv.
  
**Note**

  *The relative weight data contains only stock length individuals. This is so that I can easily compare the relative weight of fish with PSD. This is done despite the min TL being 150 mm. I may want to summarize relative weight for 150mm and greater length individuals in the future to see if young/small fish drive down or increase Wr.*

### Creating a Data File to Summarize Relative Weight by Year Length >= 150mm

```{r}
lmb.Wr <- lmb %>%
  filter(Length >= wsLMB_min) %>%
  filterD(Year>=2014)

(lmb.Wr.gcat <- Summarize(Wr~fyr*gcat,data = lmb.Wr,digits = 0) %>%
    arrange(fyr,gcat))

#1-10-2018#write.csv(lmb.Wr.gcat,file = "Data/Clean-Data/relative-weight_largemouth-bass_150.csv")
```


