---
title: "Data Preparation"
author: "Alex J. Benecke"
date: "June 15, 2017"
output: pdf_document
header-includes:
    - \setlength\parindent{24pt}
    - \usepackage{indentfirst}
editor_options: 
  chunk_output_type: console
---
```{r Load Packages, include=FALSE}
library(FSA)
library(tidyverse)
library(stringr)
library(magrittr)

```
```{r Load Data}
LMB <- read.csv("Data/Raw-Data/2016_largemouth-bass_raw.csv") %>%
  dplyr::select(FID,Site,AgeCap:Sex) %>%
  arrange(FID)
```
```{r Quick Descriptive Stats}
str(LMB) ### n = 132
tmp <- LMB %>% filter(!is.na(AgeCap))
length(tmp$AgeCap) ### Aged fish = 127, not aged = 5
mean(tmp$AgeCap)
median(tmp$AgeCap)
range(tmp$AgeCap)
table(tmp$Sex)
```
```{r Clean Data}
### Making factors factors
LMB$FID <- factor(LMB$FID)
LMB$Site <- factor(LMB$Site)
LMB$SEXCON <- factor(LMB$SEXCON)
LMB$Sex <- factor(LMB$Sex)
str(LMB)

### Removing Outliers
length(LMB$FID)
LMB[LMB$FID==55,];FID55 <- as.numeric(rownames(LMB[LMB$FID==55,])) ### Data Entry Error
#LMB[LMB$FID==89,];FID89 <- as.numeric(rownames(LMB[LMB$FID==89,])) ### Unknown Sex > ezr model fitting
rm <- LMB[c(FID55),]; rm <- filterD(rm,!is.na(FID)) ### Create df containing removed fish
LMB <- LMB[-c(FID55),] %>%
  filterD(!is.na(FID)) ### Remove outliers
length(LMB$FID)
#write.csv(LMB,file="Data/Clean-Data/2016_largemouth-bass_clean.csv")
```

Now I need to put the data into one observation per line (long) format.

```{r Put Data into Long Format, tidy=TRUE}
LMBL <- gather(LMB,Agei,Radi,Anu1:Anu9) %>%
  arrange(FID,Agei)
```

  Next I will clean up the Agei variable so that it contains only numbers, is numeric, has no NA values, and is always less than or equal to the age-at-capture.

```{r Clean LMBL, tidy=TRUE}
str_sub(LMBL$Agei,start=1,end=3) <- ""
LMBL %<>% mutate(Agei=as.numeric(Agei)) %>%
  filterD(!is.na(Radi)) %>%
  filterD(Agei<=AgeCap)
```

  Next I will perform my back-calculation using the Biological-Intercept model. Where Otoloth radius at the time of formation ($O_{0}$) is 0.035 mm (Miller and Storck 1982) and the Length of the fish at time of otolith formation ($L_{0}$) is 3.9 mm (Heang 1982).

```{r Back-Calculation, tidy=TRUE}
LMBL$BI.len <- with(LMBL,LenCap+(Radi-RadCap)*(LenCap-3.9)*(RadCap-0.035)^-1)
```
```{r Create csv File Long Format, eval=FALSE}
#write.csv(LMBL,"Data/Clean-Data/2016_largemouth-bass_long-format.csv")
```

  I have commented out the final write.csv command because I do not want to accidently remake my data file.

