---
title: "Age Bias and Precision"
author: "Alex J. Benecke"
date: "May 18, 2017"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
```{r Load Packages, include=FALSE}
library(FSA)
library(tidyr)
library(magrittr)
library(dplyr)
```
```{r Load and Prepare Data, include=TRUE}
Age <- read.csv("Data/Raw-Data/2016_largemouth-bass_age-estimates_raw.csv") %>%
  select(FID:Difference) %>%
  arrange(FID)
Age$FID <- factor(Age$FID)
length(Age$FID)
Age[Age$FID==55,];FID55 <- as.numeric(rownames(Age[Age$FID==55,]))
Age <- Age[-c(FID55),] %>%
  filterD(!is.na(FID)) ### Remove outliers
length(Age$FID)
```
```{r}
str(Age)
#write.csv(Age,file="Data/Clean-Data/2016_largemouth-bass_age-estimates_clean.csv")
```

# Comparing Age Estimates Between 3 Readers

```{r Jim and Alex Age Bias Object, include=TRUE}
ab.AB <- ageBias(AgerA~AgerB, data=Age, ref.lab = "Ager B", nref.lab = "Ager A")
```
```{r Alex and Steve Bias Object, include=TRUE}
ab.AC <- ageBias(AgerA~AgerC,data = Age,ref.lab = "Ager C", nref.lab = "Ager A")
```
```{r Jim and Steve Bias Object, include=TRUE}
ab.BC <- ageBias(AgerB~AgerC,data = Age,ref.lab = "Ager C", nref.lab = "Ager B")
```

\newpage

## Bias

 Three readers made indipendant reads of Largemouth Bass otoliths to determine the age of Largemouth Bass in our sample. Now we will look for bias between the 3 readers (readers A, B, C). We will use age bias plots to do this.

### Bias Alex (AgerA) vs Jim (AgerB) Age Est.
```{r Jim and Alex Age Bias Summary Statistics Display, echo=TRUE, tidy=TRUE}
plot(ab.AB,
     col.CIsig = "Black",
     show.range = TRUE,
     main="",
     bty="n",
     lwd = 2,
     lwd.CI = 2,
     lwd.range = 2,
     lwd.agree = 2,
     col.CI = "red",
     col.range = "red",
     cex.n = 1.25,
     cex.axis=1.5,
     yaxt = "n",
     xlab="",ylab="")
axis(2,at=seq(0,8,1),cex.axis=1.5)
mtext("Ager A",side = 2, cex = 1.5,line = 2.5)
mtext("Ager B",side = 1, cex = 1.5,line = 2.5)
mtext("Age Estimate Ager A vs. B",side = 3, cex = 1.5,line = 2.5)

### Statistics behind age bias plot
SumAB.bias <- summary(ab.AB,what = "bias")
summary(ab.AB,what = "symmetry") ### Alex's age est (A) not biased compared to Jim (B)
summary(ab.AB,what = "table")
```

### Bias Alex (AgerA) vs Steve (AgerC) Age Est.
```{r Alex and Steve Age Bias Summary Statistics Display, echo=TRUE, tidy=TRUE}
plot(ab.AC,
     col.CIsig = "Black",
     show.range = TRUE,
     main="",
     bty="n",
     lwd = 2,
     lwd.CI = 2,
     lwd.range = 2,
     lwd.agree = 2,
     col.CI = "red",
     col.range = "red",
     cex.n = 1.25,
     cex.axis=1.5,
     yaxt = "n",
     xlab="",ylab="")
axis(2,at=seq(0,8,1),cex.axis=1.5)
mtext("Ager A",side = 2, cex = 1.5,line = 2.5)
mtext("Ager C",side = 1, cex = 1.5,line = 2.5)
mtext("Age Estimate Ager A vs. C",side = 3, cex = 1.5,line = 2.5)

### Statistics behind age bias plot
SumAC.bias <- summary(ab.AC,what = "bias")
summary(ab.AC,what = "symmetry") # McNemar Suggests some age Bias between Alex and Steve
summary(ab.AC,what = "table")
```

### Bias Steve (AgerC) vs Jim (AgerB) Age Est.
```{r Jim and Steve Age Bias Summary Statistics Display, echo=TRUE, tidy=TRUE}
plot(ab.BC,col.CIsig = "Black",show.range = TRUE,main="Age Est. Ager B vs C")
```
\newpage
```{r Jim and Steve Age Bias Summary Statistics Display (Continued),, echo=TRUE, tidy=TRUE}
### Statistics behind age bias plot
summary(ab.BC,what = "bias")
summary(ab.BC,what = "symmetry") # close but no age bias between jim and steve
summary(ab.BC,what = "table")
```

## Precision

We will assess precision of age estimates using average coeficcient of variance but will also look at some other metrics.

```{r, Precision of age estimation, tidy=TRUE}
ap.ABC <- agePrecision(~AgerA+AgerB+AgerC, data=Age)

summary(ap.ABC, what = "difference")

SumABC.Prec <- summary(ap.ABC, what = "precision")
```

\newpage

| FID | AgerA (Alex) | AgerB (Jim) | AgerC (Steve) |
|:----:|:----:|:----:|:----:|
| 7 |  1 |  1 | 4  |
| 27 |  1 | 1 | 2  |
| 29 |  1 | 1  |  2 |
| 34 |  1 | 1  |  2 |
| 35  | 1  |  2 | 2  |
| 63  | 1  |  2 | 2  |
| 71  |  2 |  2 | 1  |
| 90  | 2  |  1 | 2  |
| 122 |  1 |  1 | 2  |

  There is some dissagreement about the age of some fish (FIDs 7,27,29,34,35,63,71,90,122) and some have not been aged at all (FIDs 6,28,79,81,124). I may remove these fish when fitting the growth model.
  
  Age estimates were similar between all three readers (AgerA, AgerB, and AgerC). AgerA was found to underestimate the ages of 2 year old fish when compared to AgerC (t =`r round(SumAC.bias[2,7],2)`, p = `r round(SumAC.bias[2,8],3)`) but not when compared to AgerB (t = `r round(SumAB.bias[2,7],2)`, p = `r round(SumAB.bias[2,8],3)`). We found no other significant bias in the age estimations made by the three readers. **Fish with inconsistent age estimates** ($n = 9$) **were removed from later analysis to eliminate this bias [(or) will be assigned ages based on the majority agreement between 2 readers]**. Overall, there was a high level of precision between age estimates provided by our three readers ($n =$ `r SumABC.Prec[[2]]`, `r round(SumABC.Prec[[6]],2)` \% Agreement, ACV = `r round(SumABC.Prec[[4]],2)`).













































































































