---
title: "Fitting von Bertalanffy Growth Models with Fixed Sex Effects"
author: "Alex J. Benecke"
date: "November 30, 2017"
output: pdf_document
header-includes:
    - \setlength\parindent{24pt}
    - \usepackage{indentfirst}
editor_options: 
  chunk_output_type: console
---
### Load Required Packages and Data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(FSA)
library(tidyverse)
library(nlme)
```
```{r Load and Prep Data}
LMBL <- read.csv("Data/Clean-Data/2016_largemouth-bass_long-format.csv") %>%
  select(FID:BI.len) %>%
  arrange(FID,Agei)

### Making factors factors
LMBL$FID <- factor(LMBL$FID)
LMBL$Site <- factor(LMBL$Site)
LMBL$SEXCON <- factor(LMBL$SEXCON)
LMBL$Sex <- factor(LMBL$Sex)
```
```{r Display data}
str(LMBL)
headtail(LMBL)
```

Creating the von Bertalanffy function.

```{r Build von Bertalanffy Growth Model}
LVB <- function(x, Linf, K, t0){
  y = Linf * (1 - exp(-K * (x - t0)))
  y
}
LVB <- vbFuns()

LVB(5, 422.8, 0.39, -0.40)
LVB(5, Linf = c(422.8, 0.39, -0.40)) ### Should be the same output
```

## Modeling Fixed Effects for Sex  

  **Lets look at the von Bertalanffy growth model fits with a fixed sex term on the parameter estimates.**
  
  
  In order to fit the sex model I will have to remove the individual with no sex. In order to compare the sex model to the no sex model I will have to re-fit the no sex model without the individual with unknown sex. This means making new df excluding this individual. Then, making a new grouped data object. Finnaly, I will need to  rerun the nlme function with out this fish to get the nlme.mod2 output [*This was All done in the fit-von-bertalanffy-growth-model.rmd file*].
  
#### Removing the individual with no sex (FID=89??)

```{r remove no sex individual}
### Just looking at data
head(LMBL)
### Finding fish with unknown sex
(unknown.sex <- filterD(LMBL,Sex==0))
### Getting row number for fish with the unknown sex
(FID89 <- as.numeric(row.names(LMBL[LMBL$Sex==0,])))
### removing the fish with unknown sex from the data set
length(LMBL$FID) ### just seeing the number of rows in the data set
length(unique(LMBL$FID)) ### just seeing the number of fish
LMBL <- LMBL[-c(FID89),]
length(LMBL$FID)
length(unique(LMBL$FID)) ### Good! looks like only FID 89 was removed
### Lets make sure there is no empty row in my data 
LMBL <- filterD(LMBL,!is.na(FID))
### and lets just take a quick look at the data
str(LMBL)
headtail(LMBL)
```

#### Remake Grouped Data Object
```{r Group Data Without Unknown Sex Fish}
datgr = groupedData(BI.len ~ Agei|FID, data = LMBL,
                  labels = list(x = "Age", y = "Size"),
                  units = list(x = "(Years)", y = "(mm)"))
```


### Full Sex Model, {sexmod.lkt}
**Sex terms on all model parameters** 

  I'm going to skip this for now it seems to be taking forever. I have this model output already converged, however, I did so with ML estimation instead of REML so I cannot compare with other model fits. I dont think this model should be used anyways since I'm not aware of a biological reason male and female largemouth bass would have a different $t_0$.
  
  **Failed**

```{r Full Sex Model, eval=FALSE}
sexmod.lkt <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex-1, K ~ Sex-1, t0 ~ Sex-1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647),
                                        K = c(0.4359,0.4359), 
                                        t0 = c(-0.3127,-0.3127))),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
save(sexmod.lkt, 
     file = "model-output/sexmod.lkt.rda")
```

```{r Load sexmod.lkt, include=FALSE, eval=FALSE}
load("model-output/sexmod.lkt.rda")
```
```{r Results of sexmod.lkt, echo=FALSE, eval=FALSE}
print(paste("Iterations =", sexmod.lkt$numIter))
summary(sexmod.lkt)
intervals(sexmod.lkt)
anova(sexmod.lkt)
plot(sexmod.lkt)
```

### {$L_{\infty}$, K} Sex Model, {sexmod.lk}

  I'm going to skip this for now it seems to be taking forever. I have this model output already converged, however, I did so with ML estimation instead of REML so I cannot compare with other model fits.

  Unfortunaltly, I do think this is the sex model that would make the most sense biologically except for maybe the $L_{\infty}$ of K sex models.

**Failed**
```{r L and K Sex Model, eval=FALSE, eval=FALSE}
sexmod.lk <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex-1, K ~ Sex-1, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647), 
                                        K = c(0.4359,0.4359), 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
save(sexmod.lk, 
     file = "model-output/sexmod.lk.rda")
```

### {$L_{\infty}$, $t_{0}$} Sex Model, {sexmod.lt}

```{r L and t0 Model, eval=FALSE}
sexmod.lt <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex-1, K ~ 1, t0 ~ Sex-1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647), 
                                        K = 0.4359, 
                                        t0 = c(-0.3127,-0.3127))),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.lt, 
#     file = "model-output/sexmod.lt.rda")
```

### {K, $t_{0}$}, Sex Model, {sexmod.kt}

```{r K and t0 Sex Model, eval=FALSE}
sexmod.kt <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ 1, K ~ Sex-1, t0 ~ Sex-1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = 389.3647, 
                                        K = c(0.4359,0.4359), 
                                        t0 = c(-0.3127,-0.3127))),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.kt, 
#     file = "model-output/sexmod.kt.rda")
```


### {$L_{\infty}$}, Sex Model, {sexmod.l}

```{r Linf Sex Model, eval=FALSE}
sexmod.l <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex-1, K ~ 1, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647), 
                                        K = 0.4359, 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.l, 
#     file = "model-output/sexmod.l.rda")
```

### {K}, Sex Model, {sexmod.k}

```{r K Sex Model, eval=FALSE}
sexmod.k <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ 1, K ~ Sex-1, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = 389.3647, 
                                        K = c(0.4359,0.4359), 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.k, 
#     file = "model-output/sexmod.k.rda")
```

### {$t_0$}, Sex Model, {sexmod.t}

```{r t0 Sex Model, eval=FALSE}
sexmod.t <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ 1, K ~ 1, t0 ~ Sex-1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = 389.3647, 
                                        K = c(0.4359,0.4359), 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.t, 
#     file = "model-output/sexmod.t.rda")
```


### zero intercept sex models ???

  I will now try fitting the same sex models with a zero intercept (I think this is right might be slope). this is done by removing the -1 from VBPARAM ~ Sex-1 argument.

  Rather than go through all of these right now which maybe I should I will just pick my favorits and go from there.
  
### {$L_{\infty}$, K} Intercept Sex Model {sexmod.lk.int} 

**Failed**
```{r L and K Intercept Sex Model, eval=FALSE}
sexmod.lk.int <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex, K ~ Sex, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647), 
                                        K = c(0.4359,0.4359), 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=100,
                                        pnlsMaxIter=1000,
                                        msMaxIter=100,
                                        niterEM=100))
save(sexmod.lk.int, 
     file = "model-output/sexmod.lk.int.rda")
```

### {$L_{\infty}$} Intercept Sex Model {sexmod.l.int}

```{r L Intercept Sex Model, eval=FALSE}
sexmod.l.int <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex, K ~ 1, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647), 
                                        K = 0.4359, 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=100,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.l.int, 
#     file = "model-output/sexmod.l.int.rda")
```

### {K} Intercept Sex Model {sexmod.k.int}

**Failed**
```{r K Intercept Sex Model, eval=FALSE}
sexmod.k.int <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ 1, K ~ Sex, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = 389.3647, 
                                        K = c(0.4359,0.4359), 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=100,
                                        pnlsMaxIter=1000,
                                        msMaxIter=100,
                                        niterEM=100))
save(sexmod.k.int, 
     file = "model-output/sexmod.k.int.rda")
```



### Look at Model Fits

```{r Load model fits}
load("model-output/nlme.mod2.rda")

#load("model-output/sexmod.lkt.rda")
#load("model-output/sexmod.lk.rda")
load("model-output/sexmod.lt.rda")
load("model-output/sexmod.kt.rda")
load("model-output/sexmod.l.rda")
load("model-output/sexmod.k.rda")
load("model-output/sexmod.t.rda")

#load("model-output/sexmod.lkt.int.rda")
#load("model-output/sexmod.lk.int.rda")
#load("model-output/sexmod.lt.int.rda")
#load("model-output/sexmod.kt.int.rda")
load("model-output/sexmod.l.int.rda")
#load("model-output/sexmod.k.int.rda")
#load("model-output/sexmod.t.int.rda")
```

### Results of {$L_{\infty}$}, Sex Model, {sexmod.l}

```{r Results of sexmod.l, echo=FALSE}
print(paste("Iterations =", sexmod.l$numIter))
summary(sexmod.l)
intervals(sexmod.l)
anova(sexmod.l)
plot(sexmod.l)
```
```{r sexmod.l Plot and effects}
fixef(sexmod.l)
ranef(sexmod.l)
coef(sexmod.l)[1,3]
```


```{r Plot sexmod.l}
Axes <- seq(100,450,by=50)
Years <- seq(0,10,by=1)

## plot individual fish data
## plot the fixed parameter model
## plot individual fish models

### rgb(0,1,0,0.25, maxColorValue=1)

plot(jitter(LMBL$Agei),LMBL$BI.len,
     col=ifelse(LMBL$Sex==1,rgb(0,0,1,0.25, maxColorValue=1),rgb(1,0,0,0.25, maxColorValue=1)),
     pch=19,
     ylim=c(100,500),
     xlim=c(0,10),
     xlab = "Age (Years)",
     ylab = "Back-Calculated Length (mm)",
     bty="n",
     yaxt="n",
     xaxt="n")

axis(2,at = Axes)
axis(1,at = Years)

abline(h=425.4478748,lty=2,col="blue") ### Males
abline(h=439.0623179,lty=2,col="red") ### females

x <- seq(1,11,by=1)
lines(x, fixef(sexmod.l)[1] * (1 - exp(-fixef(sexmod.l)[3] * (x - (fixef(sexmod.l)[4])))),
      lwd=3, 
      col="blue") ### Males

for(i in 1:125){
lines(x, coef(sexmod.l)[i,1] * (1 - exp(- coef(sexmod.l)[i,3]   
* (x - ( coef(sexmod.l)[i,4] )))),lwd=3,col=rgb(0,0,1,0.1),lty=3) } ### Males

lines(x, fixef(sexmod.l)[2] * (1 - exp(-fixef(sexmod.l)[3] * (x - (fixef(sexmod.l)[4])))),
      lwd=3, 
      col="red") ### Females

for(i in 1:125){
lines(x, coef(sexmod.l)[i,2] * (1 - exp(- coef(sexmod.l)[i,3]   
* (x - ( coef(sexmod.l)[i,4] )))),lwd=3,col=rgb(1,0,0,0.1),lty=3) } ### Females

legend("topleft",
       legend = print(expression(L[i]==425 %*% (1 - e **{-0.39 %*% (t[i] + 0.39)}))),
       bty="n",
       cex=1.15,
       text.col = "blue") ### VB Equation Males

legend("top",
       legend = print(expression(L[i]==439 %*% (1 - e **{-0.39 %*% (t[i] + 0.39)}))),
       bty="n",
       cex=1.15,
       text.col = "red") ### VB Equation Females

```

### {sexmod.l.int}

```{r Results of sexmod.l.int, echo=FALSE}
print(paste("Iterations =", sexmod.l.int$numIter))
summary(sexmod.l.int) ### linf.Sex2 doesn't make sense
intervals(sexmod.l.int)
anova(sexmod.l.int)
plot(sexmod.l.int)
```
```{r sexmod.l.int Plot and effects}
fixef(sexmod.l.int)
ranef(sexmod.l.int)
coef(sexmod.l.int)[1,3]
```



### Comparing Models

```{r model comparisons}
AIC(nlme.mod2,sexmod.lt,sexmod.kt,sexmod.l,sexmod.k,sexmod.t,sexmod.l.int)
BIC(nlme.mod2,sexmod.lt,sexmod.kt,sexmod.l,sexmod.k,sexmod.t,sexmod.l.int)
anova(nlme.mod2,sexmod.lt,sexmod.kt,sexmod.l,sexmod.k,sexmod.t,sexmod.l.int)

anova(nlme.mod2,sexmod.l,sexmod.l.int)
anova(sexmod.l,nlme.mod2,sexmod.l.int)

anova(sexmod.l,sexmod.k,nlme.mod2)

anova(sexmod.l,nlme.mod2)
anova(nlme.mod2,sexmod.l)

anova(sexmod.l,sexmod.l.int)
anova(sexmod.l.int,sexmod.l)
```

See sexmod.l

  There is a significant difference between the nlme.mod2 and sexmod.l with sexmod.l being the model that best describes the data (loglikliehood,df=11, L Ratio = 17.4423, p<0.0001).

