---
title: "Fitting von Bertalanffy Growth Models with Fixed Sex Effects"
author: "Alex J. Benecke"
date: "November 30, 2017"
output: pdf_document
header-includes:
    - \setlength\parindent{24pt}
    - \usepackage{indentfirst}
editor_options: 
  chunk_output_type: console
---
### Load Required Packages and Data
```{r Setup and Load Packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(FSA)
library(tidyverse)
library(nlme)
```
```{r Load and Prep Data, tidy=TRUE}
LMBL <- read.csv("Data/Clean-Data/2016_largemouth-bass_long-format.csv") %>%
  arrange(FID,Agei)

### Making factors factors
LMBL$FID <- factor(LMBL$FID)
LMBL$Site <- factor(LMBL$Site)
LMBL$SEXCON <- factor(LMBL$SEXCON)
LMBL$Sex <- factor(LMBL$Sex)
```
```{r Display Data, eval=FALSE}
str(LMBL)
headtail(LMBL)
```

Creating the von Bertalanffy function.

```{r Build von Bertalanffy Growth Model, tidy=TRUE, eval=FALSE}
LVB <- function(x, Linf, K, t0){
  y = Linf * (1 - exp(-K * (x - t0)))
  y
}
LVB <- vbFuns()

LVB(5, 422.8, 0.39, -0.40)
LVB(5, Linf = c(422.8, 0.39, -0.40)) ### Should be the same output
```

## Modeling Fixed Effects for Sex  

  **Lets look at the von Bertalanffy growth model fits with a fixed sex term on the parameter estimates.**
  
  
  In order to fit the sex model I will have to remove the individual with no sex. In order to compare the sex model to the no sex model I will have to re-fit the no sex model without the individual with unknown sex. This means making new df excluding this individual. Then, making a new grouped data object. Finally, I will need to  re-run the nlme function with out this fish to get the nlme.mod2 output.
  
#### Removing the individual with no sex (FID=89)

```{r remove no sex individual, tidy=TRUE, eval=FALSE}
### Just looking at data
head(LMBL)
### Finding fish with unknown sex
(unknown.sex <- filterD(LMBL,Sex==0))
### Getting row number for fish with the unknown sex
(FID89 <- as.numeric(row.names(LMBL[LMBL$Sex==0,])))
### removing the fish with unknown sex from the data set
length(LMBL$FID) ### just seeing the number of rows in the data set
length(unique(LMBL$FID)) ### just seeing the number of fish
LMBL <- LMBL[-c(FID89),]
length(LMBL$FID)
length(unique(LMBL$FID)) ### Good! looks like only FID 89 was removed
### Lets make sure there is no empty row in my data 
LMBL <- filterD(LMBL,!is.na(FID))
### and lets just take a quick look at the data
str(LMBL)
headtail(LMBL)
```

#### Remake Grouped Data Object
```{r Group Data Without Unknown Sex Fish}

datgr = groupedData(BI.len ~ Agei|FID, data = LMBL,
                  labels = list(x = "Age", y = "Size"),
                  units = list(x = "(Years)", y = "(mm)"))
```


### Full Sex Model, {sexmod.lkt}
**Sex terms on all model parameters** 

  I'm going to skip this for now it seems to be taking forever. I have this model output already converged, however, I did so with ML estimation instead of REML so I cannot compare with other model fits. I dont think this model should be used anyways since I'm not aware of a biological reason male and female largemouth bass would have a different $t_0$.
  
  **Failed**

```{r Full Sex Model, eval=FALSE, tidy=TRUE}
sexmod.lkt <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex-1, K ~ Sex-1, t0 ~ Sex-1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647),
                                        K = c(0.4359,0.4359), 
                                        t0 = c(-0.3127,-0.3127))),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
save(sexmod.lkt, 
     file = "model-output/sexmod.lkt.rda")
```

```{r Load sexmod.lkt, include=FALSE, eval=FALSE}
load("model-output/sexmod.lkt.rda")
```
```{r Results of sexmod.lkt, echo=FALSE, eval=FALSE}
print(paste("Iterations =", sexmod.lkt$numIter))
summary(sexmod.lkt)
intervals(sexmod.lkt)
anova(sexmod.lkt)
plot(sexmod.lkt)
```

### {$L_{\infty}$, K} Sex Model, {sexmod.lk}

  I'm going to skip this for now it seems to be taking forever. I have this model output already converged, however, I did so with ML estimation instead of REML so I cannot compare with other model fits.

  Unfortunaltly, I do think this is the sex model that would make the most sense biologically except for maybe the $L_{\infty}$ of K sex models.

**Failed**
```{r L and K Sex Model, eval=FALSE, tidy=TRUE}
sexmod.lk <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex-1, K ~ Sex-1, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647), 
                                        K = c(0.4359,0.4359), 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
save(sexmod.lk, 
     file = "model-output/sexmod.lk.rda")
```

### {$L_{\infty}$, $t_{0}$} Sex Model, {sexmod.lt}

```{r L and t0 Model, eval=FALSE, tidy=TRUE}
sexmod.lt <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex-1, K ~ 1, t0 ~ Sex-1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647), 
                                        K = 0.4359, 
                                        t0 = c(-0.3127,-0.3127))),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.lt, 
#     file = "model-output/sexmod.lt.rda")
```

### {K, $t_{0}$}, Sex Model, {sexmod.kt}

```{r K and t0 Sex Model, eval=FALSE, tidy=TRUE}
sexmod.kt <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ 1, K ~ Sex-1, t0 ~ Sex-1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = 389.3647, 
                                        K = c(0.4359,0.4359), 
                                        t0 = c(-0.3127,-0.3127))),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.kt, 
#     file = "model-output/sexmod.kt.rda")
```


### {$L_{\infty}$}, Sex Model, {sexmod.l}

```{r Linf Sex Model, eval=FALSE, tidy=TRUE}
sexmod.l <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex-1, K ~ 1, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647), 
                                        K = 0.4359, 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.l, 
#     file = "model-output/sexmod.l.rda")
```

### {K}, Sex Model, {sexmod.k}

```{r K Sex Model, eval=FALSE, tidy=TRUE}
sexmod.k <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ 1, K ~ Sex-1, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = 389.3647, 
                                        K = c(0.4359,0.4359), 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.k, 
#     file = "model-output/sexmod.k.rda")
```

### {$t_0$}, Sex Model, {sexmod.t}

```{r t0 Sex Model, eval=FALSE, tidy=TRUE}
sexmod.t <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ 1, K ~ 1, t0 ~ Sex-1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = 389.3647, 
                                        K = c(0.4359,0.4359), 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=1000,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.t, 
#     file = "model-output/sexmod.t.rda")
```

\newpage

### Zero Intercept Sex Models 

  I will now try fitting the same sex models with a zero intercept (I think this is right might be slope). this is done by removing the -1 from VBPARAM ~ Sex-1 argument.

  Rather than go through all of these right now which maybe I should I will just pick my favorits and go from there.
  
### {$L_{\infty}$, K} Intercept Sex Model {sexmod.lk.int} 

**Failed**
```{r L and K Intercept Sex Model, eval=FALSE}
sexmod.lk.int <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex, K ~ Sex, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647), 
                                        K = c(0.4359,0.4359), 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=100,
                                        pnlsMaxIter=1000,
                                        msMaxIter=100,
                                        niterEM=100))
save(sexmod.lk.int, 
     file = "model-output/sexmod.lk.int.rda")
```

### {$L_{\infty}$} Intercept Sex Model {sexmod.l.int}

```{r L Intercept Sex Model, eval=FALSE}
sexmod.l.int <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ Sex, K ~ 1, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = c(389.3647,389.3647), 
                                        K = 0.4359, 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=100,
                                        pnlsMaxIter=100,
                                        msMaxIter=100,
                                        niterEM=100))
#save(sexmod.l.int, 
#     file = "model-output/sexmod.l.int.rda")
```

### {K} Intercept Sex Model {sexmod.k.int}

**Failed**
```{r K Intercept Sex Model, eval=FALSE}
sexmod.k.int <- nlme::nlme(BI.len ~ LVB(Agei, Linf, K, t0), data = datgr,
                            fixed = list(Linf ~ 1, K ~ Sex, t0 ~ 1),
                            random = Linf+K+t0 ~ 1,
                            start = list(fixed = 
                                      c(Linf = 389.3647, 
                                        K = c(0.4359,0.4359), 
                                        t0 = -0.3127)),
                           method= "REML",
                           control=list(opt="nlminb",
                                        maxIter=100,
                                        pnlsMaxIter=1000,
                                        msMaxIter=100,
                                        niterEM=100))
save(sexmod.k.int, 
     file = "model-output/sexmod.k.int.rda")
```



### Look at Model Fits

```{r Load Model Output, tidy=TRUE}
### The outputs commented out failed to converge
load("model-output/nlme.mod2.rda")

#load("model-output/sexmod.lkt.rda")
#load("model-output/sexmod.lk.rda")
load("model-output/sexmod.lt.rda")
load("model-output/sexmod.kt.rda")
load("model-output/sexmod.l.rda")
load("model-output/sexmod.k.rda")
load("model-output/sexmod.t.rda")

#load("model-output/sexmod.lkt.int.rda")
#load("model-output/sexmod.lk.int.rda")
#load("model-output/sexmod.lt.int.rda")
#load("model-output/sexmod.kt.int.rda")
load("model-output/sexmod.l.int.rda")
#load("model-output/sexmod.k.int.rda")
#load("model-output/sexmod.t.int.rda")
```

### Results of {$L_{\infty}$}, Sex Model, {sexmod.l}

```{r Results of sexmod.l, echo=FALSE, tidy=TRUE}
print(paste("Iterations =", sexmod.l$numIter))
summary(sexmod.l)
intervals(sexmod.l)
anova(sexmod.l)
plot(sexmod.l)
```
```{r Try to test difference between male and female Linf, include=FALSE}
### Try to test difference between male and female Linf
sexmod.l$coefficients$fixed[c(1:2)]
sexmod.l$coefficients$random$FID[,1]
coef(sexmod.l)[,c(1:2)]
```
```{r sexmod.l Plot and effects, include=FALSE}
fixef(sexmod.l)
ranef(sexmod.l)
coef(sexmod.l)[1,3]
```


```{r Plot sexmod.l, echo=FALSE, tidy=TRUE}
Axes <- seq(100,450,by=50)
Years <- seq(0,10,by=1)

## plot individual fish data
## plot the fixed parameter model
## plot individual fish models

### rgb(0,1,0,0.25, maxColorValue=1)

plot(jitter(LMBL$Agei),LMBL$BI.len,
     col=ifelse(LMBL$Sex==1,rgb(0,0,1,0.25, maxColorValue=1),rgb(1,0,0,0.25, maxColorValue=1)),
     pch=19,
     ylim=c(100,500),
     xlim=c(0,10),
     xlab = "Age (Years)",
     ylab = "Back-Calculated Length (mm)",
     bty="n",
     yaxt="n",
     xaxt="n")

axis(2,at = Axes)
axis(1,at = Years)

abline(h=425.4478748,lty=2,col="blue") ### Males
abline(h=439.0623179,lty=2,col="red") ### females

x <- seq(1,11,by=1)
lines(x, fixef(sexmod.l)[1] * (1 - exp(-fixef(sexmod.l)[3] * (x - (fixef(sexmod.l)[4])))),
      lwd=3, 
      col="blue") ### Males

for(i in 1:125){
lines(x, coef(sexmod.l)[i,1] * (1 - exp(- coef(sexmod.l)[i,3]   
* (x - ( coef(sexmod.l)[i,4] )))),lwd=3,col=rgb(0,0,1,0.1),lty=3) } ### Males

lines(x, fixef(sexmod.l)[2] * (1 - exp(-fixef(sexmod.l)[3] * (x - (fixef(sexmod.l)[4])))),
      lwd=3, 
      col="red") ### Females

for(i in 1:125){
lines(x, coef(sexmod.l)[i,2] * (1 - exp(- coef(sexmod.l)[i,3]   
* (x - ( coef(sexmod.l)[i,4] )))),lwd=3,col=rgb(1,0,0,0.1),lty=3) } ### Females

legend("topleft",
       legend = print(expression(L[i]==425 %*% (1 - e **{-0.39 %*% (t[i] + 0.39)}))),
       bty="n",
       cex=1.15,
       text.col = "blue") ### VB Equation Males

legend("topright",
       legend = print(expression(L[i]==439 %*% (1 - e **{-0.39 %*% (t[i] + 0.39)}))),
       bty="n",
       cex=1.15,
       text.col = "red") ### VB Equation Females

legend("bottomright",
       legend = c("Male", "Female"),
       col=c("blue", "red"),
       lty=1, pch=19, bty="n")
```

### $L_{\infty}$ Intercept Sex Model {sexmod.l.int}

```{r Results of sexmod.l.int, echo=FALSE}
print(paste("Iterations =", sexmod.l.int$numIter))
summary(sexmod.l.int) ### linf.Sex2 doesn't make sense
intervals(sexmod.l.int)
anova(sexmod.l.int)
plot(sexmod.l.int)
```
```{r sexmod.l.int Plot and effects, eval=FALSE}
fixef(sexmod.l.int)
ranef(sexmod.l.int)
coef(sexmod.l.int)[1,3]
```

\newpage

### Comparing Models

```{r model comparisons, tidy=TRUE}
AIC(nlme.mod2,sexmod.lt,sexmod.kt,sexmod.l,sexmod.k,sexmod.t,sexmod.l.int)
BIC(nlme.mod2,sexmod.lt,sexmod.kt,sexmod.l,sexmod.k,sexmod.t,sexmod.l.int)
anova(nlme.mod2,sexmod.lt,sexmod.kt,sexmod.l,sexmod.k,sexmod.t,sexmod.l.int)

anova(nlme.mod2,sexmod.l,sexmod.l.int)
anova(sexmod.l,nlme.mod2,sexmod.l.int)

anova(sexmod.l,sexmod.k,nlme.mod2)
```


### {$L_{\infty}$} Sex Model vs. No Sex Model (sexmod.l vs. nlme.mod2)

```{r Compare sexmod.l and nlme.mod2, tidy=TRUE}
anova(sexmod.l,nlme.mod2)
anova(nlme.mod2,sexmod.l)
```

  There is a significant difference between the nlme.mod2 and sexmod.l with sexmod.l being the model that best describes the data (loglikliehood,df=11, L Ratio = 17.4423, p<0.0001).

### {$L_{\infty}$} Sex Model vs. {$L_{\infty}$} Intercept Sex Model (sexmod.l vs. sexmod.l.int)

```{r Compare sexmod.l and sexmod.l.int, tidy=TRUE}
anova(sexmod.l,sexmod.l.int)
anova(sexmod.l.int,sexmod.l)
```


```{r Summarize sexmod.l For Results, include=FALSE}
anova(nlme.mod2,sexmod.l)
summary(sexmod.l)
intervals(sexmod.l)
```


  Estimates for the von Bertalanffy growth parameters were obtained from 336 obsercations of 125 largemouth bass. Two additional fish were removed prior to model fitting due to inaccurate or incomplete data. The von Bertalanffy growth model fit with a sex term on the $L_{\infty}$ described the data better than when no sex term was used (Likelihood Ratio Test, Liklihood Ratio = 17, df = 11, p < 0.0001). The maximum asymptotic length ($L_{\infty}$) was larger for female largemouth bass ($L_{\infty}$ = 439, SE = 8.55) than male largemouth bass ($L_{\infty}$ = 425, SE = 8.43) ($F_{2, 208}$ = 20992, p < 0.001). The brody growth rate coefficient (K) and the theoretical time at which length was zero ($t_{0}$) were fit without sex effects as 0.39 (SE = 0.015) and -0.39 (SE = 0.044) respectively. 


```{r Quick Descriptive Stats, include=FALSE}
### Number observations for Male LMB
tmp.m <- LMBL %>% filterD(Sex==1)
levels(tmp.m$Sex)
str(tmp.m)
nrow(tmp.m) ### Number of Observations = 179
length(unique(tmp.m$FID)) ### Number of individuals = 56

### Number observations for Female LMB
tmp.f <- LMBL %>% filterD(Sex==2)
levels(tmp.f$Sex)
str(tmp.f)
nrow(tmp.f) ### Number of Observations = 157
length(unique(tmp.f$FID)) ### Number of individuals = 69
```
  
  
| Sex    | Number of Individuals | Number of Observations |
|:------:|:---------------------:|:----------------------:|
| Male   |         56            |          179           |
| Female |         69            |          157           |
| Total  |         125           |          336           |
